---
title: "Metanalaysis for Down Syndrome"

output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
    collapsed: true
    smooth_scroll: true
    number_sections: false
---

```{r, echo=FALSE, warning=FALSE, comment=NA, message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, comment=NA, message=FALSE, width=1000, cache=TRUE, cache.lazy=FALSE, dpi=300, dev=c('png', 'pdf'))
options(width=1200)
```


```{r wdirectory}
# setwd("~/OneDrive - CRG - Centre de Regulacio Genomica/Dropbox (CRG)/metanalysis")
load("201911_de_list_ds.RData")

```


```{r, subset_metanalysis}

#function to select for each contract the top DEGs
subset_metanalysis=function(list_array, adjpval=0.1, max_n_genes=20000, abslog2FC=0){
  #list_array is a list of list with names "list_ensembl", "list_adjpval", "list_log2FC"
  
  indexes1=lapply(list_array$adjpval, FUN=function(x) which(x < adjpval))
  nomi=names(list_array$names)
  list_array$names= lapply(1:length(list_array$names), FUN=function(x) (list_array$names[[x]][(indexes1[[x]])]))
  list_array$adjpval= lapply(1:length(list_array$adjpval), FUN=function(x) (list_array$adjpval[[x]][(indexes1[[x]])]))
  list_array$log2FC= lapply(1:length(list_array$log2FC), FUN=function(x) (list_array$log2FC[[x]][(indexes1[[x]])]))
  names(list_array$names)=nomi
  names(list_array$adjpval)=nomi
  names(list_array$log2FC)=nomi
  
  
  subset=which(lapply(list_array$names, length)>0)
  list_array$names=list_array$names[subset]
  list_array$adjpval=list_array$adjpval[subset]
  list_array$log2FC=list_array$log2FC[subset]
  nomi=names(list_array$names)
  
  indexes1=lapply(list_array$log2FC, FUN=function(x) which((x > abslog2FC )| (x < -abslog2FC )))
  
  list_array$names= lapply(1:length(list_array$names), FUN=function(x) (list_array$names[[x]][(indexes1[[x]])]))
  list_array$adjpval= lapply(1:length(list_array$adjpval), FUN=function(x) (list_array$adjpval[[x]][(indexes1[[x]])]))
  list_array$log2FC= lapply(1:length(list_array$log2FC), FUN=function(x) (list_array$log2FC[[x]][(indexes1[[x]])]))
  names(list_array$names)=nomi
  names(list_array$adjpval)=nomi
  names(list_array$log2FC)=nomi
  
  subset=which(lapply(list_array$names, length)>0)
  list_array$names=list_array$names[subset]
  list_array$adjpval=list_array$adjpval[subset]
  list_array$log2FC=list_array$log2FC[subset]
  
  
  for(x in 1:length(list_array$adjpval)){
    if(!is.na(list_array$adjpval[[x]][max_n_genes])){
      max_genes=max_n_genes
      a=TRUE
      while(a){
        if(list_array$adjpval[[x]][max_genes] == list_array$adjpval[[x]][max_genes+1]){
          max_genes=max_genes+1
          
        } else{a=FALSE}}
      
      list_array$names[[x]]= list_array$names[[x]][1:max_genes][!is.na(list_array$names[[x]][1:max_genes])]
      list_array$adjpval[[x]]=list_array$adjpval[[x]][1:max_genes][!is.na(list_array$adjpval[[x]][1:max_genes])]
      list_array$log2FC[[x]]=list_array$log2FC[[x]][1:max_genes][!is.na(list_array$log2FC[[x]][1:max_genes])]
      
    }
  }
  return(list_array) 
}
```


```{r create_input}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))
```


```{r find_foldchanges}
find_fold_changes=function(ensgene, dataset=list_array){
  dataset_true=dataset
  dataset_true$names=dataset$names[unlist(lapply(dataset$names, FUN=function(x) (ensgene %in% x)))]
  dataset_true$adjpval=dataset$adjpval[unlist(lapply(dataset$names, FUN=function(x) (ensgene %in% x)))]
  dataset_true$log2FC=dataset$log2FC[unlist(lapply(dataset$names, FUN=function(x) (ensgene %in% x)))]
  
  
  output=unlist(lapply(1:length(dataset_true$log2FC), FUN=function(x)(dataset_true$log2FC[[x]][which(dataset_true$names[[x]]==ensgene)])))
  names(output)=names(dataset_true$names)
  return(output)
}

```

```{r}
# consider as DEGs the top500 with at least 1.5 fold change and adjusted p-value < 0.05
list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


```


### Checking distribution of occurrences

```{r}
library(poweRlaw)
m1 = displ$new(table(unlist(list_array.05_fc1.5_max500$names)))
m1$setPars(estimate_pars(m1))

m2 = dispois$new(table(unlist(list_array.05_fc1.5_max500$names)))
m2$setPars(estimate_pars(m2))

m3 = disexp$new(table(unlist(list_array.05_fc1.5_max500$names)))
m3$setPars(estimate_pars(m3))


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))

paste("power-law versus Poisson:", compare_distributions(m1, m2)$p_one_sided)
paste("power-law versus Exponential:",compare_distributions(m1, m3)$p_one_sided)
paste("power-law versus log-normal:",compare_distributions(m1, m4)$p_one_sided)
paste("poisson versus log-normal:",compare_distributions(m2, m4)$p_one_sided)
paste("exponential versus log-normal:",compare_distributions(m3, m4)$p_one_sided)

hist(table(unlist(list_array.05_fc1.5_max500$names)), xlab="number of comparisons where the gene is found DE", main="", breaks=20 )
abline(v=4, col="red", lty=2)
plot(m1, ylab="Cumulative distribution function", xlab="number of occurences")
lines(m1, lty=2)
lines(m2, col=2, lty=2)
lines(m3, col=3, lty=2)
lines(m4, col=4, lty=2)

legend("bottomleft", c("power-law", "Poisson", "Exponential","log-normal"), fill=c(1,2,3,4), cex=.65, bty="n")

hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))

```


### Distribution tissue/species

```{r, tissuespecies, fig.width=7, fig.height=10}

summary=read.csv("summary_metanalysis_DS.csv",  header=T)
rownames(summary)=summary$Name
summary_sub=summary[names(list_array.05_fc1.5_max500$names),]



# library(randomcoloR)
# n <- 16
# palette <- distinctColorPalette(n)
vec=paste(summary_sub$Sample_type, summary_sub$Species, sep=".")
vec=gsub("Homo sapiens", "human", vec)
vec=gsub("Mus musculus", "mouse", vec)

vec.human=vec[grep("human", vec)]
vec.mouse=vec[grep("mouse", vec)]


dat=as.data.frame(table(vec.mouse))
rownames(dat)=gsub(".mouse", "", as.character(dat$vec.mouse))

dat2=as.data.frame(table(vec.human))
dat$tissue=rownames(dat)
rownames(dat2)=gsub(".human", "", as.character(dat2$vec.human))
dat2$tissue=rownames(dat2)



dat_m=merge(dat, dat2, by="tissue", all=TRUE)

rownames(dat_m)=dat_m$tissue

dat_m=dat_m[,c(3,5)]
dat_m[is.na(dat_m)] <- 0
colnames(dat_m)=c("mouse", "human")

par(mar=c(5,15,4,1))
barplot(t(dat_m[names(sort(table(summary_sub$Sample_type))[which(sort(table(summary_sub$Sample_type))>0)]),]), horiz=T, las=2, col=c("red","black"), main="Sample Type")
legend("right", c("mouse", "human"), fill=c("red", "black"), bty="n")


```


```{r}


# par(mar=c(5,27,4,2))
# barplot(sort(table(summary_sub$Sample_type)), horiz=T, las=2, col=rainbow(26), main="Sample Type")


par(mar=c(5,27,4,2))

barplot(sort(table(summary_sub$Platform)), horiz=T, las=2, col=rainbow(29), main="Platform")

par(mar=c(5,27,4,2))

barplot(sort(table(summary_sub$Model)[!is.na(table(summary_sub$Model))]), horiz=T, las=2, col=rainbow(9), main="Mouse Model")
```

```{r}
gene_frequency=sort(table(unlist(list_array.05_fc1.5_max500$names)), decreasing=T)


gene_frequency_top=gene_frequency[which(gene_frequency>=gene_frequency[round(length(gene_frequency)*0.05)])]
gene_frequency_top_genes=unique(unlist(lapply(names(gene_frequency_top), function(x)(strsplit(x, split=".", fixed=T)))))


gene_matrix=data.frame(names=gene_frequency_top_genes, occurences=as.numeric(gene_frequency[gene_frequency_top_genes]))

```

### Comparison with previous metanalysis

```{r}
previous=read.csv("ds_metanalysis.csv", header=T)

common_name=c("GSE6283_amnio",
"GSE6283_cvs",

"GSE1294",
"GSE1294b",
"GSE1611",
"GSE5390",
"EMEXP409",	

"GSE1397.crb",
"GSE1397.crbl",
"GSE1397.astro",
"GSE1397.heart",
"GSE9762",
"GSE16176")


common_geo=c("GSE6283",
"GSE1294",
"GSE1611",
"GSE5390",
"E-MEXP-409",	
"GSE1397",
"GSE9762",
"GSE16176")

```


```{r}
par(mar=c(5,4,4,2))

library(tagcloud)

tags=rownames(summary)
colori=rep("black", length(rownames(summary)))
colori[which(tags %in% common_name)]="red"

set.seed(1)
tagcloud(tags, col=colori)


tags2=names(table(summary$GEO.ArrayExpress))

colori2=rep("black", length(tags2))
colori2[which(tags2 %in% common_geo)]="red"

set.seed(1)
tagcloud(tags2, weights =table(summary$GEO.ArrayExpress),  col=colori2)
```


### Analysis per tissue

```{r}
brain_studies= which(summary$macro_type=="Brain")
undifferentiated_studies= which(summary$macro_type=="Undifferentiated")
fibroblasts_studies= which(summary$macro_type=="Fibroblasts")
placenta_amnios_studies= which(summary$macro_type=="Placenta/amnios")
heart_studies= which(summary$macro_type=="Heart")
muscle_studies= which(summary$macro_type=="Muscle")
blood_studies= which(summary$macro_type=="Blood/bone marrow")
liver_studies= which(summary$macro_type=="Liver")
sperm_studies= which(summary$macro_type=="Sperm Cells")
endothelia_studies= which(summary$macro_type=="Endothelia")
```

#### Brain

```{r}
list_arrayALL=list_array

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[brain_studies]

list_array$adjpval=list_array$adjpval[brain_studies]

list_array$log2FC=list_array$log2FC[brain_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.06)])]))



DS_brain_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`


#### blood_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[blood_studies]

list_array$adjpval=list_array$adjpval[blood_studies]

list_array$log2FC=list_array$log2FC[blood_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_blood_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`


#### undifferentiated_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[undifferentiated_studies]

list_array$adjpval=list_array$adjpval[undifferentiated_studies]

list_array$log2FC=list_array$log2FC[undifferentiated_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_undifferentiated_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`



#### fibroblasts_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[fibroblasts_studies]

list_array$adjpval=list_array$adjpval[fibroblasts_studies]

list_array$log2FC=list_array$log2FC[fibroblasts_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_fibroblasts_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`


#### placenta_amnios_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[placenta_amnios_studies]

list_array$adjpval=list_array$adjpval[placenta_amnios_studies]

list_array$log2FC=list_array$log2FC[placenta_amnios_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_placenta_amnios_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`



#### heart_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[heart_studies]

list_array$adjpval=list_array$adjpval[heart_studies]

list_array$log2FC=list_array$log2FC[heart_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_heart_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`




#### muscle_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[muscle_studies]

list_array$adjpval=list_array$adjpval[muscle_studies]

list_array$log2FC=list_array$log2FC[muscle_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_muscle_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`



#### liver_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[liver_studies]

list_array$adjpval=list_array$adjpval[liver_studies]

list_array$log2FC=list_array$log2FC[liver_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_liver_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`


#### sperm_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[sperm_studies]

list_array$adjpval=list_array$adjpval[sperm_studies]

list_array$log2FC=list_array$log2FC[sperm_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_sperm_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`



#### endothelia_studies

```{r}

list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))


list_array$names=list_array$names[endothelia_studies]

list_array$adjpval=list_array$adjpval[endothelia_studies]

list_array$log2FC=list_array$log2FC[endothelia_studies]

```

`r length(list_array$log2FC)` contrasts remaining

```{r}

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )


for(i in 1:length(list_array.05_fc1.5_max500$names)){
  
  list_array.05_fc1.5_max500$names[[i]]=list_array.05_fc1.5_max500$names[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$adjpval[[i]]=list_array.05_fc1.5_max500$adjpval[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
  list_array.05_fc1.5_max500$log2FC[[i]]=list_array.05_fc1.5_max500$log2FC[[i]][which(!duplicated(list_array.05_fc1.5_max500$names[[i]]))]
}


m4 = dislnorm$new(table(unlist(list_array.05_fc1.5_max500$names)))
m4$setPars(estimate_pars(m4))
pvalues.05_fc1.5_max500=rev(dist_cdf(m4, lower_tail = FALSE))


hubs.05_fc1.5_max500=unique(names(table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[which(pvalues.05_fc1.5_max500<0.05)])]))



DS_endothelia_studies=hubs.05_fc1.5_max500
```

minimum dataset to be consistently DE: `r table(unlist(list_array.05_fc1.5_max500$names))[names(pvalues.05_fc1.5_max500[length(hubs.05_fc1.5_max500)])]`






```{r}
DS_tissue_genes=list(DS_brain_studies, DS_undifferentiated_studies, DS_fibroblasts_studies, DS_placenta_amnios_studies, DS_heart_studies, DS_muscle_studies,DS_blood_studies, DS_liver_studies, DS_sperm_studies, DS_endothelia_studies)


names(DS_tissue_genes)=c("brain", "iPSCs/ESCs", "fibroblasts", "placenta/amnios","heart", "muscle", "blood", "liver","sperm cells", "endothelia")

str(DS_tissue_genes)

```
```{r}
create_overlap_matrix3=function(lista){
  mat1a=matrix(rep(NA,length(lista)*length(lista)), length(lista),length(lista)) #number
  mat=matrix(rep(NA,length(lista)*length(lista)), length(lista),length(lista)) #overlap coef
  rownames(mat)=names(lista)
  colnames(mat)=names(lista)
  rownames(mat1a)=names(lista)
  colnames(mat1a)=names(lista)
  for (o in 1:length(lista)){
    for (n in 1:length(lista)){
  
      op=as.character(lista[[o]])
      np=as.character(lista[[n]])
      
      mat[o,n]=length(intersect(op,np))/min(length(op), length(np))
      mat1a[o,n]=length(intersect(op,np))
      
    }}

  return(list(mat1a, mat))
}
```

```{r}
DS_tissue_genes=DS_tissue_genes[which(sapply(DS_tissue_genes, length)>0)]

DS_overlaps=create_overlap_matrix3(DS_tissue_genes)

rownames(DS_overlaps[[2]])=paste0(names(DS_tissue_genes), " (",sapply(DS_tissue_genes, length), ")" )
```

### Tissue-macro category specific DE gene


```{r, fig.width=8, fig.height=8}
library(gplots)
library(RColorBrewer)
heatmap.2(DS_overlaps[[2]], dendrogram="none", scale="none", density.info="none", trace="none", col=colorRampPalette(brewer.pal(9, "OrRd"))(256), main="", key.title="", Rowv=FALSE, Colv=rev("Rowv"), cex.axis=.5, cexRow=1.1, cexCol=1.1,cellnote=DS_overlaps[[1]], keysize=1,key.xlab ="overlap coefficient",srtCol=45, margins=c(8,12))
```


```{r}
library(biomaRt)
human = useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl",
mirror = "useast")
genes=unique(unlist(DS_tissue_genes))
genesfromtissues=getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "chromosome_name"),filters="ensembl_gene_id", values=genes, mart = human)
# save(DS_tissue_genes, file="DS_tissue_genes.rda")
# ada
# 
# genesfromtissues$hgnc_symbol[which(genesfromtissues$hgnc_symbol=="")]=c("Lnc-SSU72P8-2","Lnc-GPM6A-1", "Lnc-EFR3A-1", "lnc-COX10-3","TEC", "Lnc-ATXN1-8", "ALDOA")

rownames(genesfromtissues)=genesfromtissues$ensembl_gene_id

```




```{r}
mat_graph2=cbind(rep(NA, sum(sapply(DS_tissue_genes, length))), rep(NA, sum(sapply(DS_tissue_genes, length))))

index=0
for(i in 1:length(DS_tissue_genes)){
  mat_graph2[,1][(index+1):(index+length(DS_tissue_genes[[i]]))]=names(DS_tissue_genes)[[i]]
  for(j in 1:length(DS_tissue_genes[[i]])){
    mat_graph2[,2][index+j]=DS_tissue_genes[[i]][j]
  }
  index=index+length(DS_tissue_genes[[i]])
}



library(igraph)
# graph_datasets=graph_from_adjacency_matrix(DS_overlaps[[1]], mode="undirected", weighted=TRUE, diag=FALSE)

```



```{r}

library(igraph)
graph_coex2=graph_from_data_frame(as.data.frame(mat_graph2), directed = FALSE, vertices = NULL)


V(graph_coex2)$size =5

 V(graph_coex2)$size[grep("ENSG", V(graph_coex2)$name)]=2


# summary=read.csv("summary_metanalysis_DS.csv",  header=T)
# rownames(summary)=summary$Name
# macro_type_col=as.factor(as.character(summary[V(graph_coex2)$name[1:86],]$macro_type))
# study_labels=as.character(summary[V(graph_coex2)$name[1:86],]$Sample_type)
# 
# labels=rep(NA, vcount(graph_coex2))
# 
# 
# labels[1:86]=study_labels

V(graph_coex2)$color="white"
library(RColorBrewer)

V(graph_coex2)$color[1:length(DS_tissue_genes)]=brewer.pal(length(DS_tissue_genes), "Set1")[1:length(DS_tissue_genes)]

# V(graph_coex2)$color[which(V(graph_coex2)$name %in% V(graph_coex)$name)]="turquoise"

entropicgenes=names(which(table(unlist(DS_tissue_genes))>1))


genesfromtissues=genesfromtissues[V(graph_coex2)$name,]


genesfromtissues$chromosome_name[which(genesfromtissues$chromosome_name=="CHR_HSCHR16_1_CTG1")]="16"
genesfromtissues$chromosome_name[which(genesfromtissues$chromosome_name=="CHR_HSCHR21_3_CTG1_1")]="21"
genesfromtissues$chromosome_name[which(genesfromtissues$chromosome_name=="CHR_HSCHR21_4_CTG1_1")]="21"
genesfromtissues$chromosome_name[which(genesfromtissues$chromosome_name=="CHR_HSCHR21_5_CTG2")]="21"

V(graph_coex2)$shape =rep("circle", vcount(graph_coex2))
V(graph_coex2)$shape[which(genesfromtissues$chromosome_name=="21")]="square"


labels=genesfromtissues$hgnc_symbol
labels[1:length(DS_tissue_genes)]=names(DS_tissue_genes)
col_labels=rep("red", vcount(graph_coex2))
col_labels[1:length(DS_tissue_genes)]="black"


size_labels=rep(.75, vcount(graph_coex2))
size_labels[1:length(DS_tissue_genes)]=1




df_sizes=as.data.frame(cbind(gene_matrix[,1], log2(gene_matrix[,2])))

V(graph_coex2)$size[which(V(graph_coex2)$name %in% df_sizes$V1 )]=as.numeric(as.character(df_sizes$V2))[which(df_sizes$V1 %in% V(graph_coex2)$name)]

labels2=rep(NA,length(labels))
labels2[1:6]=labels[1:6]

V(graph_coex2)$color[which(V(graph_coex2)$name %in% entropicgenes)]="black"


labels2[which(V(graph_coex2)$name %in% entropicgenes)]=labels[which(V(graph_coex2)$name %in% entropicgenes)]






# coords2=tk_coords(virtualplot)
# save(coords2, file="coords2.rda")

load("coords2.rda")

coords2=coords2[c(1:6,which(V(graph_coex2)$name %in% gene_matrix[,1])),]
todelete=setdiff(V(graph_coex2)$name, gene_matrix[,1])

todelete=setdiff(todelete, todelete[1:6])
graph_coex2=delete_vertices(graph_coex2, todelete)
par(mar=c(0,0,0,5),xpd=TRUE)
set.seed(1)
plot(graph_coex2,vertex.label=labels2[c(1:6,which(V(graph_coex2)$name %in% gene_matrix[,1]))], vertex.label.color=col_labels[c(1:6,which(V(graph_coex2)$name %in% gene_matrix[,1]))],vertex.label.cex=size_labels[c(1:6,which(V(graph_coex2)$name %in% gene_matrix[,1]))], layout=coords2)


legend("topleft", 
       c("chr21", "other chrs"), bty="n",cex=1,
       pch=c(0,1), title="deregulated in:")

legend("bottom", 
       c("low", "high"), bty="n",pt.cex=c(1,3),
       pch=1, title="log2(occurences):",cex=1)



```


```{r}
list_array=list(list_ensembl, list_adjpval, list_log2FC)

names(list_array)=c("names", "adjpval", "log2FC")

## correct names
list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)]=unlist(lapply(list_array$names$EMTAB312[grep(" /// ",list_array$names$EMTAB312)], FUN=function(x)(strsplit(x, " /// ")[[1]][1])))
list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )

list_array.05_fc1.5_max500=subset_metanalysis(list_array, adjpval = 0.05, abslog2FC = log2(1.5), max_n_genes = 500 )

```

## Chromosome distribution



```{r}
library(biomaRt)
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl",host="oct2018.archive.ensembl.org")

interesting_genes=getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "chromosome_name", "band"),filters="ensembl_gene_id", values=gene_matrix$names, mart = human)
interesting_genes$chromosome_name[which(interesting_genes$chromosome_name=="CHR_HSCHR14_7_CTG1")]=14

interesting_genes$chromosome_name[grep("CHR_HSCHR", interesting_genes$chromosome_name)]=21


interesting_genes$description=unlist(lapply(interesting_genes$description, FUN=function(x)strsplit(x, split='[', fixed=T)[[1]][1]))


rownames(interesting_genes)=interesting_genes$ensembl_gene_id

interesting_genes=interesting_genes[gene_matrix$names,]



interesting_genes$occurences=gene_matrix$occurences

interesting_genes$combinations=gene_matrix$combinations




interesting_genes$median_log2FC=rep(NA, dim(interesting_genes)[1])
interesting_genes$sd_log2FC=rep(NA, dim(interesting_genes)[1])
interesting_genes$mean_log2FC=rep(NA, dim(interesting_genes)[1])


for(i in 1:dim(interesting_genes)[1]){
  ensgene=interesting_genes$ensembl_gene_id[i]
  fcs=find_fold_changes(ensgene, list_array.05_fc1.5_max500)
  interesting_genes$median_log2FC[i]=median(fcs)
  interesting_genes$sd_log2FC[i]=sd(fcs)
  interesting_genes$mean_log2FC[i]=mean(fcs)
  
}



dtchr=aggregate(median_log2FC~chromosome_name, FUN=median, data=interesting_genes)
prova=dtchr[1:22,]
prova$chromosome_name=as.numeric(prova$chromosome_name)
dtchr=rbind(prova[order(prova$chromosome_name),], dtchr[dim(dtchr)[1],])
nomi=dtchr$chromosome_name
dtchr=dtchr[,2]
names(dtchr)=nomi


```

```{r}
par(mar=c(12,4,4,2))
barplot(table(interesting_genes$chromosome_name)[c(1,12,16:22,2:11, 13:15, length(table(interesting_genes$chromosome_name)))], las=2, xlab="chromosome", ylab = "number of genes", main="consistently DE genes")
```

```{r}

all_genes=getBM(attributes=c("ensembl_gene_id", "chromosome_name"),values="*", mart = human)
all_genes_table=table(all_genes$chromosome_name)

table1=table(interesting_genes$chromosome_name)[c(1,12,16:22,2:11, 13:15, length(table(interesting_genes$chromosome_name)))]
par(mar=c(12,4,4,2))
barplot(table1/all_genes_table[names(table1)], las=2, xlab="chromosome", ylab = "% of genes", main="consistently DE genes")
```




```{r}
interesting_genes$chr_band=as.factor(paste(interesting_genes$chromosome_name, interesting_genes$band, sep=""))

median_chrband=split(interesting_genes$median_log2FC, interesting_genes$chr_band)

median_chrband=median_chrband[c(grep("^1p",names(median_chrband)),grep("^1q",names(median_chrband)),
grep("^2p",names(median_chrband)),grep("^2q",names(median_chrband)),
grep("^3p",names(median_chrband)),grep("^3q",names(median_chrband)),
grep("^4p",names(median_chrband)),grep("^4q",names(median_chrband)),
grep("^5p",names(median_chrband)),grep("^5q",names(median_chrband)),
grep("^6p",names(median_chrband)),grep("^6q",names(median_chrband)),
grep("^7p",names(median_chrband)),grep("^7q",names(median_chrband)),
grep("^8p",names(median_chrband)),grep("^8q",names(median_chrband)),
grep("^9p",names(median_chrband)),grep("^9q",names(median_chrband)),
grep("^10p",names(median_chrband)),grep("^10q",names(median_chrband)),
grep("^11p",names(median_chrband)),grep("^11q",names(median_chrband)),
grep("^12p",names(median_chrband)),grep("^12q",names(median_chrband)),
grep("^13p",names(median_chrband)),grep("^13q",names(median_chrband)),
grep("^14p",names(median_chrband)),grep("^14q",names(median_chrband)),
grep("^15p",names(median_chrband)),grep("^15q",names(median_chrband)),
grep("^16p",names(median_chrband)),grep("^16q",names(median_chrband)),
grep("^17p",names(median_chrband)),grep("^17q",names(median_chrband)),
grep("^18p",names(median_chrband)),grep("^18q",names(median_chrband)),
grep("^19p",names(median_chrband)),grep("^19q",names(median_chrband)),
grep("^20p",names(median_chrband)),grep("^20q",names(median_chrband)),
grep("^21p",names(median_chrband)),grep("^21q",names(median_chrband)),
grep("^22p",names(median_chrband)),grep("^22q",names(median_chrband)),
grep("^Xp",names(median_chrband)),grep("Xq",names(median_chrband)))]




tol23rainbow= c("#ffffff", "#771155", "#AA4488", "#CC99BB", "#114477", "#4477AA", "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#117744", "#44AA77", "#88CCAA", "#777711", "#AAAA44", "#DDDD77", "#774411", "#AA7744", "#DDAA77", "#771122", "#AA4455", "#DD7788", "#000000")

colors=c(rep(tol23rainbow[1],length(c(grep("^1p",names(median_chrband)),grep("^1q",names(median_chrband))))),
         rep(tol23rainbow[2],length(c(grep("^2p",names(median_chrband)),grep("^2q",names(median_chrband))))),
rep(tol23rainbow[3],length(c(grep("^3p",names(median_chrband)),grep("^3q",names(median_chrband))))),
rep(tol23rainbow[4],length(c(grep("^4p",names(median_chrband)),grep("^4q",names(median_chrband))))),
rep(tol23rainbow[5],length(c(grep("^5p",names(median_chrband)),grep("^5q",names(median_chrband))))),
rep(tol23rainbow[6],length(c(grep("^6p",names(median_chrband)),grep("^6q",names(median_chrband))))),
rep(tol23rainbow[7],length(c(grep("^7p",names(median_chrband)),grep("^7q",names(median_chrband))))),
rep(tol23rainbow[8],length(c(grep("^8p",names(median_chrband)),grep("^8q",names(median_chrband))))),
rep(tol23rainbow[9],length(c(grep("^9p",names(median_chrband)),grep("^9q",names(median_chrband))))),
rep(tol23rainbow[10],length(c(grep("^10p",names(median_chrband)),grep("^10q",names(median_chrband))))),
rep(tol23rainbow[11],length(c(grep("^11p",names(median_chrband)),grep("^11q",names(median_chrband))))),
rep(tol23rainbow[12],length(c(grep("^12p",names(median_chrband)),grep("^12q",names(median_chrband))))),
rep(tol23rainbow[13],length(c(grep("^13p",names(median_chrband)),grep("^13q",names(median_chrband))))),
rep(tol23rainbow[14],length(c(grep("^14p",names(median_chrband)),grep("^14q",names(median_chrband))))),
rep(tol23rainbow[15],length(c(grep("^15p",names(median_chrband)),grep("^15q",names(median_chrband))))),
rep(tol23rainbow[16],length(c(grep("^16p",names(median_chrband)),grep("^16q",names(median_chrband))))),
rep(tol23rainbow[17],length(c(grep("^17p",names(median_chrband)),grep("^17q",names(median_chrband))))),
rep(tol23rainbow[18],length(c(grep("^18p",names(median_chrband)),grep("^18q",names(median_chrband))))),
rep(tol23rainbow[19],length(c(grep("^19p",names(median_chrband)),grep("^19q",names(median_chrband))))),
rep(tol23rainbow[20],length(c(grep("^20p",names(median_chrband)),grep("^20q",names(median_chrband))))),
rep(tol23rainbow[21],length(c(grep("^21p",names(median_chrband)),grep("^21q",names(median_chrband))))),
rep(tol23rainbow[22],length(c(grep("^22p",names(median_chrband)),grep("^22q",names(median_chrband))))),
rep(tol23rainbow[23],length(c(grep("^Xp",names(median_chrband)),grep("Xq",names(median_chrband))))))


```


```{r, fig.width=70, fig.height=7}

boxplot(median_chrband, xlab="region", ylab="log2FC",col=colors,las=2)
abline(h=0, lty="dashed", col="gray")
abline(h=log2(1.5), lty="dashed", col="red")
abline(h=-log2(1.5), lty="dashed", col="blue")
```



### % of genes on band

```{r}
table_chrband=table(interesting_genes$chr_band)

table_chrband=table_chrband[c(grep("^1p",names(table_chrband)),grep("^1q",names(table_chrband)),
grep("^2p",names(table_chrband)),grep("^2q",names(table_chrband)),
grep("^3p",names(table_chrband)),grep("^3q",names(table_chrband)),
grep("^4p",names(table_chrband)),grep("^4q",names(table_chrband)),
grep("^5p",names(table_chrband)),grep("^5q",names(table_chrband)),
grep("^6p",names(table_chrband)),grep("^6q",names(table_chrband)),
grep("^7p",names(table_chrband)),grep("^7q",names(table_chrband)),
grep("^8p",names(table_chrband)),grep("^8q",names(table_chrband)),
grep("^9p",names(table_chrband)),grep("^9q",names(table_chrband)),
grep("^10p",names(table_chrband)),grep("^10q",names(table_chrband)),
grep("^11p",names(table_chrband)),grep("^11q",names(table_chrband)),
grep("^12p",names(table_chrband)),grep("^12q",names(table_chrband)),
grep("^13p",names(table_chrband)),grep("^13q",names(table_chrband)),
grep("^14p",names(table_chrband)),grep("^14q",names(table_chrband)),
grep("^15p",names(table_chrband)),grep("^15q",names(table_chrband)),
grep("^16p",names(table_chrband)),grep("^16q",names(table_chrband)),
grep("^17p",names(table_chrband)),grep("^17q",names(table_chrband)),
grep("^18p",names(table_chrband)),grep("^18q",names(table_chrband)),
grep("^19p",names(table_chrband)),grep("^19q",names(table_chrband)),
grep("^20p",names(table_chrband)),grep("^20q",names(table_chrband)),
grep("^21p",names(table_chrband)),grep("^21q",names(table_chrband)),
grep("^22p",names(table_chrband)),grep("^22q",names(table_chrband)),
grep("^Xp",names(table_chrband)),grep("Xq",names(table_chrband)))]



colors2=c(rep(tol23rainbow[1],length(c(grep("^1p",names(table_chrband)),grep("^1q",names(table_chrband))))),
         rep(tol23rainbow[2],length(c(grep("^2p",names(table_chrband)),grep("^2q",names(table_chrband))))),
rep(tol23rainbow[3],length(c(grep("^3p",names(table_chrband)),grep("^3q",names(table_chrband))))),
rep(tol23rainbow[4],length(c(grep("^4p",names(table_chrband)),grep("^4q",names(table_chrband))))),
rep(tol23rainbow[5],length(c(grep("^5p",names(table_chrband)),grep("^5q",names(table_chrband))))),
rep(tol23rainbow[6],length(c(grep("^6p",names(table_chrband)),grep("^6q",names(table_chrband))))),
rep(tol23rainbow[7],length(c(grep("^7p",names(table_chrband)),grep("^7q",names(table_chrband))))),
rep(tol23rainbow[8],length(c(grep("^8p",names(table_chrband)),grep("^8q",names(table_chrband))))),
rep(tol23rainbow[9],length(c(grep("^9p",names(table_chrband)),grep("^9q",names(table_chrband))))),
rep(tol23rainbow[10],length(c(grep("^10p",names(table_chrband)),grep("^10q",names(table_chrband))))),
rep(tol23rainbow[11],length(c(grep("^11p",names(table_chrband)),grep("^11q",names(table_chrband))))),
rep(tol23rainbow[12],length(c(grep("^12p",names(table_chrband)),grep("^12q",names(table_chrband))))),
rep(tol23rainbow[13],length(c(grep("^13p",names(table_chrband)),grep("^13q",names(table_chrband))))),
rep(tol23rainbow[14],length(c(grep("^14p",names(table_chrband)),grep("^14q",names(table_chrband))))),
rep(tol23rainbow[15],length(c(grep("^15p",names(table_chrband)),grep("^15q",names(table_chrband))))),
rep(tol23rainbow[16],length(c(grep("^16p",names(table_chrband)),grep("^16q",names(table_chrband))))),
rep(tol23rainbow[17],length(c(grep("^17p",names(table_chrband)),grep("^17q",names(table_chrband))))),
rep(tol23rainbow[18],length(c(grep("^18p",names(table_chrband)),grep("^18q",names(table_chrband))))),
rep(tol23rainbow[19],length(c(grep("^19p",names(table_chrband)),grep("^19q",names(table_chrband))))),
rep(tol23rainbow[20],length(c(grep("^20p",names(table_chrband)),grep("^20q",names(table_chrband))))),
rep(tol23rainbow[21],length(c(grep("^21p",names(table_chrband)),grep("^21q",names(table_chrband))))),
rep(tol23rainbow[22],length(c(grep("^22p",names(table_chrband)),grep("^22q",names(table_chrband))))),
rep(tol23rainbow[23],length(c(grep("^Xp",names(table_chrband)),grep("Xq",names(table_chrband))))))


```




```{r, fig.width=120, fig.height=7}
all_genes=getBM(attributes=c("ensembl_gene_id", "chromosome_name", "band"),values="*", mart = human)


all_genes$chr_band=as.factor(paste(all_genes$chromosome_name, all_genes$band, sep=""))


all_genes_table=table(all_genes$chr_band)



all_genes_table=all_genes_table[c(grep("^1p",names(all_genes_table)),grep("^1q",names(all_genes_table)),
grep("^2p",names(all_genes_table)),grep("^2q",names(all_genes_table)),
grep("^3p",names(all_genes_table)),grep("^3q",names(all_genes_table)),
grep("^4p",names(all_genes_table)),grep("^4q",names(all_genes_table)),
grep("^5p",names(all_genes_table)),grep("^5q",names(all_genes_table)),
grep("^6p",names(all_genes_table)),grep("^6q",names(all_genes_table)),
grep("^7p",names(all_genes_table)),grep("^7q",names(all_genes_table)),
grep("^8p",names(all_genes_table)),grep("^8q",names(all_genes_table)),
grep("^9p",names(all_genes_table)),grep("^9q",names(all_genes_table)),
grep("^10p",names(all_genes_table)),grep("^10q",names(all_genes_table)),
grep("^11p",names(all_genes_table)),grep("^11q",names(all_genes_table)),
grep("^12p",names(all_genes_table)),grep("^12q",names(all_genes_table)),
grep("^13p",names(all_genes_table)),grep("^13q",names(all_genes_table)),
grep("^14p",names(all_genes_table)),grep("^14q",names(all_genes_table)),
grep("^15p",names(all_genes_table)),grep("^15q",names(all_genes_table)),
grep("^16p",names(all_genes_table)),grep("^16q",names(all_genes_table)),
grep("^17p",names(all_genes_table)),grep("^17q",names(all_genes_table)),
grep("^18p",names(all_genes_table)),grep("^18q",names(all_genes_table)),
grep("^19p",names(all_genes_table)),grep("^19q",names(all_genes_table)),
grep("^20p",names(all_genes_table)),grep("^20q",names(all_genes_table)),
grep("^21p",names(all_genes_table)),grep("^21q",names(all_genes_table)),
grep("^22p",names(all_genes_table)),grep("^22q",names(all_genes_table)),
grep("^Xp",names(all_genes_table)),grep("Xq",names(all_genes_table)))]


table_chrband2=rep(0, length(all_genes_table))
names(table_chrband2)  =names(all_genes_table)

table_chrband2[names(table_chrband)]=table_chrband

barplot(table_chrband2/all_genes_table, xlab="region", ylab="% of genes",col=colors2, las=2)


pdf(height=7, width = 120, file="Percent_gene_band.pdf")
barplot(table_chrband2/all_genes_table, xlab="region", ylab="% of genes",col=colors2, las=2)
abline(h=.1, col="red", lty=2)
dev.off()
```






### Entropy

```{r}
sectors=cbind(names(list_array.05_fc1.5_max500$names),as.character(summary[names(list_array.05_fc1.5_max500$names),"Sample_type"]),as.character(summary[names(list_array.05_fc1.5_max500$names),"macro_type"]))
colnames(sectors)=c("study", "sample_type", "macro_type")
sectors=as.data.frame(sectors)
rownames(sectors)=sectors$study

list_sample_type=rep(list(NA), length(levels(as.factor(sectors$sample_type))))
names(list_sample_type)=levels(as.factor(sectors$sample_type))
for(i in 1:length(list_sample_type)){
  studies=as.character(unique(sectors$study[which(sectors$sample_type == names(list_sample_type)[i])]))
  list_sample_type[[i]]=unique(unlist(list_array.05_fc1.5_max500$names[studies]))
}


list_macro_type=rep(list(NA), length(levels(as.factor(sectors$macro_type))))
names(list_macro_type)=levels(as.factor(sectors$macro_type))
for(i in 1:length(list_macro_type)){
  studies=as.character(unique(sectors$study[which(sectors$macro_type == names(list_macro_type)[i])]))
  list_macro_type[[i]]=unique(unlist(list_array.05_fc1.5_max500$names[studies]))
} 
```

```{r interestinggenes2bc}

interesting_genes$Blood=rep(0, dim(interesting_genes)[1])
interesting_genes$Blood[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`Blood/bone marrow`)]=1

interesting_genes$Cerebellum=rep(0, dim(interesting_genes)[1])
interesting_genes$Cerebellum[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Cerebellum)]=1

interesting_genes$Cortex=rep(0, dim(interesting_genes)[1])
interesting_genes$Cortex[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Cortex)]=1


interesting_genes$Endothelia=rep(0, dim(interesting_genes)[1])
interesting_genes$Endothelia[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Endothelia)]=1

interesting_genes$ESCs=rep(0, dim(interesting_genes)[1])
interesting_genes$ESCs[which(interesting_genes$ensembl_gene_id %in% list_sample_type$ESCs)]=1

interesting_genes$fetalliver=rep(0, dim(interesting_genes)[1])
interesting_genes$fetalliver[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`Fetal liver`)]=1

interesting_genes$Fibroblasts=rep(0, dim(interesting_genes)[1])
interesting_genes$Fibroblasts[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Fibroblasts)]=1

interesting_genes$Forebrain=rep(0, dim(interesting_genes)[1])
interesting_genes$Forebrain[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Forebrain)]=1



interesting_genes$Heart=rep(0, dim(interesting_genes)[1])
interesting_genes$Heart[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Heart)]=1

interesting_genes$Hippocampus=rep(0, dim(interesting_genes)[1])
interesting_genes$Hippocampus[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Hippocampus)]=1

interesting_genes$iPSCs=rep(0, dim(interesting_genes)[1])
interesting_genes$iPSCs[which(interesting_genes$ensembl_gene_id %in% list_sample_type$iPSCs)]=1


interesting_genes$Liver=rep(0, dim(interesting_genes)[1])
interesting_genes$Liver[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Liver)]=1



interesting_genes$Muscle=rep(0, dim(interesting_genes)[1])
interesting_genes$Muscle[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Muscle)]=1

interesting_genes$NPCs=rep(0, dim(interesting_genes)[1])
interesting_genes$NPCs[which(interesting_genes$ensembl_gene_id %in% list_sample_type$NPCs)]=1
                    
interesting_genes$Placenta=rep(0, dim(interesting_genes)[1])
interesting_genes$Placenta[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`Placenta/amnios`)]=1

# interesting_genes$Sperm=rep(0, dim(interesting_genes)[1])
# interesting_genes$Sperm[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`Sperm Cells`)]=1

interesting_genes$Thymus=rep(0, dim(interesting_genes)[1])
interesting_genes$Thymus[which(interesting_genes$ensembl_gene_id %in% list_sample_type$Thymus)]=1


interesting_genes$wholebrain=rep(0, dim(interesting_genes)[1])
interesting_genes$wholebrain[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`whole brain`)]=1

interesting_genes$wholeembryo=rep(0, dim(interesting_genes)[1])
interesting_genes$wholeembryo[which(interesting_genes$ensembl_gene_id %in% list_sample_type$`whole embryo`)]=1

mat_interesting_genes=interesting_genes[, (dim(interesting_genes)[2]-17):dim(interesting_genes)[2]]

interesting_genes$entropy=rowSums(mat_interesting_genes)

```

```{r tissuedis, fig.width = 16, fig.height=16}
library(ggplot2)
library(gplots)
library(RColorBrewer)
hm=heatmap.2(as.matrix(mat_interesting_genes), scale="none", density.info="none", trace="none", col=colorRampPalette(brewer.pal(9, "OrRd"))(256), key.title="", cex.axis=.5, cexRow=.9, 
             cexCol=.9, keysize=.75,margins=c(16,16),srtCol=45,lhei=c(1,8),
             ColSideColors =rainbow(10)[as.numeric(as.factor(c("Blood", "Brain", "Brain", "Endhotelia","Undifferentiated", "Liver", "Fibroblasts", "Brain", "Heart", "Brain", "Undifferentiated", "Liver", "Muscle", "Undifferentiated", "Placenta", "Blood", "Brain", "Undifferentiated")))])

```


```{r}
interesting_genes_macro=interesting_genes[,c(1:10)]

list_chr=split(interesting_genes_macro$ensembl_gene_id, interesting_genes_macro$chromosome_name)

list_chr=list_chr[c(as.character(1:22), "X")]


interesting_genes_macro$Blood=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Blood[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$`Blood/bone marrow`)]=1


interesting_genes_macro$Brain=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Brain[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Brain)]=1


interesting_genes_macro$Endothelia=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Endothelia[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Endothelia)]=1


interesting_genes_macro$Fibroblasts=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Fibroblasts[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Fibroblasts)]=1

interesting_genes_macro$Heart=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Heart[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Heart)]=1

interesting_genes_macro$Liver=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Liver[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Liver)]=1


interesting_genes_macro$Muscle=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Muscle[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Muscle)]=1


interesting_genes_macro$placenta=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$placenta[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$`Placenta/amnios`)]=1

interesting_genes_macro$Undifferentiated=rep(0, dim(interesting_genes_macro)[1])
interesting_genes_macro$Undifferentiated[which(interesting_genes_macro$ensembl_gene_id %in% list_macro_type$Undifferentiated)]=1



mat_interesting_genes_macro=interesting_genes_macro[, (dim(interesting_genes_macro)[2]-8):dim(interesting_genes_macro)[2]]

```

```{r tissue distribution, fig.width = 16, fig.height=16}

hm=heatmap.2(as.matrix(mat_interesting_genes_macro), scale="none", density.info="none", trace="none", col=colorRampPalette(brewer.pal(9, "OrRd"))(256), key.title="", cex.axis=.5, cexRow=.9, 
             cexCol=.9, keysize=.75,margins=c(16,16),srtCol=45,lhei=c(1,8),
             ColSideColors =rainbow(10)[as.numeric(as.factor(colnames(mat_interesting_genes_macro)))])

```


```{r}

colors=rep("white",dim(interesting_genes_macro)[1])

tissue=rep("others",dim(interesting_genes_macro)[1])

for(i in 1:6){
  colors[which(interesting_genes_macro$ensembl_gene_id %in% DS_tissue_genes[[which(names(DS_tissue_genes)==V(graph_coex2)$name[i])]])]=V(graph_coex2)$color[i]
  
  tissue[which(interesting_genes_macro$ensembl_gene_id %in% DS_tissue_genes[[which(names(DS_tissue_genes)==V(graph_coex2)$name[i])]])]=V(graph_coex2)$name[i]
}

colors[which(interesting_genes_macro$ensembl_gene_id %in% entropicgenes)]="black"
tissue[which(interesting_genes_macro$ensembl_gene_id %in% entropicgenes)]="entropic gene"

interesting_genes_macro$tissue=tissue



interesting_genes_macro$entropy=rowSums(mat_interesting_genes_macro)

library(ggplot2)

df=data.frame("Entropy"=interesting_genes_macro$entropy, "Group"=factor(interesting_genes_macro$tissue, levels=c("iPSCs/ESCs", "fibroblasts", "placenta/amnios", "brain", "blood", "entropic gene", "others")))
rownames(df)=interesting_genes_macro$ensembl_gene_id
set.seed(1)
p <- ggplot(df, aes(x=Group, y=Entropy)) +  geom_jitter(shape=16, position=position_jitter(0.2), col=colors)+ylab("Entropy (Macro-categories")  + stat_summary(fun.y = median, fun.min = median, fun.max = median,
                 geom = "crossbar", width = 0.5)+theme(axis.text.x = element_text(angle=45))
# Box plot with jittered points
# 0.2 : degree of jitter in x direction
p



# df2=data.frame("Entropy"=interesting_genes_macro[V(graph_coex)$name,"entropy"], "Group"=as.factor(V(graph_coex)$color))
# rownames(df2)=V(graph_coex)$name 
# 
# p <- ggplot(df2, aes(x=Group, y=Entropy))  + geom_jitter(shape=16, position=position_jitter(0.2), col=V(graph_coex)$color)+ylab("Entropy (Macro-categories") +
#     stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
#                  geom = "crossbar", width = 0.5) + scale_x_discrete(labels=c("iPSCs/ESCs", "fibroblasts", "placenta/amnios", "brain", "blood", "consistenty in >1", "consistently only disregarding tissue"))+theme(axis.text.x = element_text(angle=45))
# 
# 
# # Box plot with jittered points
# # 0.2 : degree of jitter in x direction
# p


```

### Relation entropy and HSA21


```{r}

percent_21=c(length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==1)))/length(which(interesting_genes_macro$entropy==1)),

length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==2)))/length(which(interesting_genes_macro$entropy==2)),

length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==3)))/length(which(interesting_genes_macro$entropy==3)),

length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==4)))/length(which(interesting_genes_macro$entropy==4)),

length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==5)))/length(which(interesting_genes_macro$entropy==5)),

length(intersect(which(interesting_genes_macro$chromosome_name =="21"),which(interesting_genes_macro$entropy==6)))/length(which(interesting_genes_macro$entropy==6)))




names(percent_21)=1:6
barplot(percent_21*100, xlab="entropy", ylab="% of HSA21 genes")

abline(h=length(which(interesting_genes_macro$chromosome_name =="21"))/dim(interesting_genes_macro)[1]
*100, lty="dashed", col="blue")
```

## Tissue-specific genes

TissueEnrich defines tissue-specific genes using RNA-Seq data from the HPA, GTEx, and mouse ENCODE. In order to make the tissue-specific gene calculations more robust, we only used tissues that had 2 biological replicates. The datasets used in the tool are:

HPA Dataset: RNA-Seq data across 35 human tissues (Uhln et al. 2015).
GTEx Dataset: RNA-Seq data across 29 human tissues (Ardlie et al. 2015).
Mouse ENCODE Dataset: RNA-Seq data across 17 mouse tissues (Shen et al. 2012).


Tissue-specific genes are defined using the algorithm from the HPA (Uhln et al. 2015), and can be grouped as follows:

Tissue Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to all other tissues.
Group Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a group of 2-7 tissues compared to all other tissues, and that are not considered Tissue Enriched.
Tissue Enhanced: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to the average levels in all other tissues, and that are not considered Tissue Enriched or Group Enriched.

```{r, fig.height=10}
library(TissueEnrich)

gs<-GeneSet(geneIds=interesting_genes$ensembl_gene_id,organism="Homo Sapiens",geneIdType=ENSEMBLIdentifier())

output<-teEnrichment(inputGenes = gs, tissueSpecificGeneType=2)
output2<-teEnrichment(inputGenes = gs, rnaSeqDataset = 2)
output3<-teEnrichment(inputGenes = gs, rnaSeqDataset = 3)


tissuetable=as.data.frame(assay(output[[1]]))

tissuetable3=as.data.frame(assay(output3[[1]]))

count_tissuespecificgenes=function(output){
  listgenes=c()
  for(i in 1:length(output[[2]])){
    
    if(dim(output[[2]][[i]])[1]>0){
      listgenes=append(listgenes, rownames(output[[2]][[i]]))
    }
  }
  return(unique(listgenes))
}

list_splitted=split(interesting_genes_macro, interesting_genes_macro$entropy)


percent_specific=rep(NA, length(list_splitted))
for(i in 1:length(list_splitted)){
 
  gs<-GeneSet(geneIds=rownames(list_splitted[[i]]),organism="Homo Sapiens",geneIdType=ENSEMBLIdentifier())

  output<-teEnrichment(inputGenes = gs) 

  percent_specific[i]=length(count_tissuespecificgenes(output))/dim(list_splitted[[i]])[1]
}

# > percent_specific
# [1] 0.3728814 0.2857143 0.2678571 0.1666667 0.0000000 0.0000000


interesting_genes_macro21=interesting_genes_macro[which(interesting_genes_macro$chromosome_name=="21"),]




list_splitted21=split(interesting_genes_macro21, interesting_genes_macro21$entropy)


percent_specific21=rep(NA, length(list_splitted21))
for(i in 1:length(list_splitted21)){
 
  gs<-GeneSet(geneIds=rownames(list_splitted21[[i]]),organism="Homo Sapiens",geneIdType=ENSEMBLIdentifier())

  output<-teEnrichment(inputGenes = gs) 

  percent_specific21[i]=length(count_tissuespecificgenes(output))/dim(list_splitted21[[i]])[1]
}


# > percent_specific21
# [1] 0.66666667 0.35000000 0.07894737 0.03703704 0.00000000 0.00000000

interesting_genes_macroNON21=interesting_genes_macro[which(interesting_genes_macro$chromosome_name!="21"),]




list_splittedNON21=split(interesting_genes_macroNON21, interesting_genes_macroNON21$entropy)


percent_specificNON21=rep(NA, length(list_splitted))
for(i in 1:length(list_splittedNON21)){
 
  gs<-GeneSet(geneIds=rownames(list_splittedNON21[[i]]),organism="Homo Sapiens",geneIdType=ENSEMBLIdentifier())

  output<-teEnrichment(inputGenes = gs) 

  percent_specificNON21[i]=length(count_tissuespecificgenes(output))/dim(list_splittedNON21[[i]])[1]
}
names(percent_specific)=1:6
names(percent_specific21)=1:6
names(percent_specificNON21)=1:6
par(mfrow=c(3,1), mar=c(3,4,2,2))

barplot(percent_specific*100, xlab="entropy", ylab="% tissue-specific genes", main="All genes", ylim=c(0,65))
barplot(percent_specific21*100, xlab="entropy", ylab="% tissue-specific genes", main="HSA21 genes", ylim=c(0,65))
barplot(percent_specificNON21*100, xlab="entropy", ylab="% tissue-specific genes", main="non-HSA21 genes", ylim=c(0,65))

```

```{r}
#save.image("test.rda")
```



### HSA21 neighbors

```{r}
library(STRINGdb)

string_db2 <- STRINGdb$new(version="11",species=9606,score_threshold=900, input_directory="" )


list_ds=as.data.frame(all_genes$ensembl_gene_id[grep("21", all_genes$chromosome_name)])
colnames(list_ds)="list_ds"



mapped <- string_db2$map( list_ds, "list_ds", removeUnmappedRows = FALSE )


neighbors=c()
for (i in 1:length(mapped$STRING_id[!(is.na(mapped$STRING_id))])) {
  tryCatch({
    neighbors=append(neighbors, string_db2$get_neighbors(mapped$STRING_id[!(is.na(mapped$STRING_id))][i] ))
  }, error=function(e){})
}

neighbors=unique(neighbors)




neighbors_hsa21=getBM(attributes=c("ensembl_peptide_id", "hgnc_symbol"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors, FUN=function(x)(substring(x,6)))), mart = human)

neighbors_ens=unique(getBM(attributes=c("ensembl_peptide_id", "ensembl_gene_id"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors, FUN=function(x)(substring(x,6)))), mart = human)$ensembl_gene_id)


neighbors_hsa21=unique(neighbors_hsa21$hgnc_symbol)

```


```{r,fig.width=4.8, fig.height=5.6}
detected_genes=unique((unlist(list_array.05_fc1.5_max500$names)))

set.seed(1)

sample=rownames(interesting_genes)
universe=detected_genes

value=neighbors_ens
nperm=1000


permutation_vec=rep(list(NA), nperm)
set.seed=1
for(i in 1:nperm){
  random=sample(universe, length(sample))
  permutation_vec[[i]]=length(intersect(random, value))
}

pvalue=(sum(unlist(permutation_vec )> length(intersect(value,sample)))+1)/(nperm+1)

boxplot(list(length(intersect(neighbors_ens, rownames(interesting_genes))),unlist(permutation_vec)), ylab="number of co-DE genes", xaxt="n", col=c("blue", "red"))
axis(1, c(1,2),c("HSA21 Neighbors", "1000 permutations"))
```


## co-DE gene clusters

We detected two groups of 13 genes, each of them consistently presents in 4 datasets.

```{r loadcomb}
load("~/OneDrive - CRG - Centre de Regulacio Genomica/Dropbox (CRG)/metanalysis/combinazioni/top_genes.RData")
load("~/OneDrive - CRG - Centre de Regulacio Genomica/Dropbox (CRG)/metanalysis/combinazioni/top.RData")
load("./combinazioni/all_comb13_rank.RData")
load("./combinazioni/all_comb11_rank.RData")
lista1=unlist(strsplit(names(all_comb13_rank_top)[[1]], split=".", fixed = TRUE))
lista2=unlist(strsplit(names(all_comb13_rank_top)[[2]], split=".", fixed = TRUE))

gene_matrix$combinations=rep(1, length(gene_matrix$names))

gene_matrix$combinations[which(gene_matrix$names %in% all_comb2_rank_top_genes)]=2
gene_matrix$combinations[which(gene_matrix$names %in% all_comb3_rank_top_genes)]=3
gene_matrix$combinations[which(gene_matrix$names %in% all_comb4_rank_top_genes)]=4
gene_matrix$combinations[which(gene_matrix$names %in% all_comb5_rank_top_genes)]=5
gene_matrix$combinations[which(gene_matrix$names %in% all_comb6_rank_top_genes)]=6
gene_matrix$combinations[which(gene_matrix$names %in% all_comb7_rank_top_genes)]=7
gene_matrix$combinations[which(gene_matrix$names %in% all_comb8_rank_top_genes)]=8
gene_matrix$combinations[which(gene_matrix$names %in% all_comb9_rank_top_genes)]=9
gene_matrix$combinations[which(gene_matrix$names %in% all_comb10_rank_top_genes)]=10
gene_matrix$combinations[which(gene_matrix$names %in% all_comb11_rank_top_genes)]=11
gene_matrix$combinations[which(gene_matrix$names %in% all_comb12_rank_top_genes)]=12
gene_matrix$combinations[which(gene_matrix$names %in% all_comb13_rank_top_genes)]=13


rownames(gene_matrix)=gene_matrix$names
lista1_bm=getBM(attributes=c( "hgnc_symbol", "ensembl_gene_id"),filters="ensembl_gene_id", values=lista1, mart = human)
lista2_bm=getBM(attributes=c( "hgnc_symbol", "ensembl_gene_id"),filters="ensembl_gene_id", values=lista2, mart = human)
lista3_bm=getBM(attributes=c( "hgnc_symbol", "ensembl_gene_id"),filters="ensembl_gene_id", values=rownames(gene_matrix)[which(gene_matrix$combinations==11)], mart = human)

rownames(lista1_bm)=lista1_bm$ensembl_gene_id
rownames(lista2_bm)=lista2_bm$ensembl_gene_id
rownames(lista3_bm)=lista3_bm$ensembl_gene_id

```

One was detected in the following contrasts:

```{r firstgroupdatabases}
library(kableExtra)

kable(summary[names(all_comb13dot[grep(names(sort(table(all_comb13dot), decreasing=T)[1]), all_comb13dot)]),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 

```
And included the following genes:

```{r firstgroupgenes}

kable(interesting_genes[unlist(strsplit(names(sort(table(all_comb13dot), decreasing=T)[1]), split=".", fixed = TRUE)),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T)


```

The other was found in the following contrasts:

```{r secondgroupdatabases}
kable(summary[names(all_comb13dot[grep(names(sort(table(all_comb13dot), decreasing=T)[2]), all_comb13dot)]),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 

```

And included the following genes:

```{r secondgroupgenes}
kable(interesting_genes[unlist(strsplit(names(sort(table(all_comb13dot), decreasing=T)[2]), split=".", fixed = TRUE)),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T)
```


Finally we also detected a cluster of 11 genes:

```{r 11dot}
kable(interesting_genes[rownames(lista3_bm),], "html") %>%
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T)


```

Found in the following contrasts:

```{r string3}
string3=paste(sort(rownames(lista3_bm))[1], sort(rownames(lista3_bm))[2],sort(rownames(lista3_bm))[3],
      sort(rownames(lista3_bm))[4], sort(rownames(lista3_bm))[5],sort(rownames(lista3_bm))[6],
      sort(rownames(lista3_bm))[7], sort(rownames(lista3_bm))[8],sort(rownames(lista3_bm))[9],
      sort(rownames(lista3_bm))[10], sort(rownames(lista3_bm))[11],sep=".")



kable(summary[names(all_comb11dot[ grep(string3, all_comb11dot)]),], "html") %>%
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 




```


`r intersect(interesting_genes[unlist(strsplit(names(sort(table(all_comb13dot), decreasing=T)[1]), split=".", fixed = TRUE)),]$hgnc_symbol, interesting_genes[unlist(strsplit(names(sort(table(all_comb13dot), decreasing=T)[2]), split=".", fixed = TRUE)),]$hgnc_symbol)` was in common in the two groups.




```{r}
load("./combinazioni/all_comb2_rank_top_genes.RData")

load("./combinazioni/all_comb2_rank.RData")
```





## DS network

```{r}
list_coex=lapply(names(all_comb2_rank), function(x)(strsplit(x, split=".", fixed=T)[[1]]))
mat_coex=do.call(rbind, list_coex)


library(igraph)
graph_coex=graph_from_data_frame(as.data.frame(mat_coex), directed = FALSE, vertices = NULL)


V(graph_coex)$size <- log2(gene_matrix[V(graph_coex)$name,2])

```

```{r}
library(plot3D)
# whiteblue=colorRampPalette(c("white", "blue"))
# 
# whitegray=colorRampPalette(c("yellow", "orange", "red",  "brown"))
# 
# 
# cutcolors=whiteblue(13)[gene_matrix[V(graph_coex)$name,3]]
nomi_vertex=rep(NA, vcount(graph_coex))

nomi_vertex[V(graph_coex)$name %in% lista1_bm$ensembl_gene_id]= lista1_bm[( V(graph_coex)$name[V(graph_coex)$name %in% lista1_bm$ensembl_gene_id]),1]

nomi_vertex[V(graph_coex)$name %in% lista2_bm$ensembl_gene_id]= lista2_bm[( V(graph_coex)$name[V(graph_coex)$name %in% lista2_bm$ensembl_gene_id]),1]

nomi_vertex[V(graph_coex)$name %in% lista3_bm$ensembl_gene_id]= lista3_bm[( V(graph_coex)$name[V(graph_coex)$name %in% lista3_bm$ensembl_gene_id]),1]



nomi_col=rep(NA, vcount(graph_coex))

nomi_col[V(graph_coex)$name %in% lista1_bm$ensembl_gene_id]= "red"

nomi_col[V(graph_coex)$name %in% lista2_bm$ensembl_gene_id]= "red"

nomi_col[V(graph_coex)$name %in% lista3_bm$ensembl_gene_id]= "red"



E(graph_coex)$weight=all_comb2_rank
E(graph_coex)$width=log2(E(graph_coex)$weight)


library(arrangements)
library(viridis)

edges_names=paste(mat_coex[,1], mat_coex[,2], sep=".")

edge_color=rep("transparent", ecount(graph_coex))

edge_weights=rep(1, ecount(graph_coex))

edge_color[which(edges_names %in% names(all_comb2_rank_top))]=rev(viridis(12))[1]
edge_weights[which(edges_names %in% names(all_comb2_rank_top))]=2


for(i in 1:length(all_comb3_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb3_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[2]
  
  edge_weights[which(edges_names %in% toadd_comb_dot)]=3
}


for(i in 1:length(all_comb4_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb4_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[3]
     edge_weights[which(edges_names %in% toadd_comb_dot)]=4

}

for(i in 1:length(all_comb5_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb5_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[4]
     edge_weights[which(edges_names %in% toadd_comb_dot)]=5

}


for(i in 1:length(all_comb6_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb6_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[5]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=6

}


for(i in 1:length(all_comb7_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb7_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[6]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=7

}


for(i in 1:length(all_comb8_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb8_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[7]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=8

}

for(i in 1:length(all_comb9_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb9_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[8]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=9

}

for(i in 1:length(all_comb10_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb10_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[9]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=10

}

for(i in 1:length(all_comb11_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb11_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[10]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=11

}

for(i in 1:length(all_comb12_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb12_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[11]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=12

}




for(i in 1:length(all_comb13_rank_top)){
  toadd_names=unique(unlist(strsplit(names(all_comb13_rank_top)[[i]], split=".", fixed = TRUE)))
   toadd_comb=combinations(toadd_names,2, layout="list")
  toadd_comb=lapply(toadd_comb,sort)
  
  toadd_comb_dot=sapply(toadd_comb, function(x)(paste(x[1], x[2],sep=".")))
  
  edge_color[which(edges_names %in% toadd_comb_dot)]=rev(viridis(12))[12]
  edge_weights[which(edges_names %in% toadd_comb_dot)]=13

}


V(graph_coex)$shape <- ifelse((interesting_genes[V(graph_coex)$name,"chromosome_name"]=="21"), "square", "circle")

edge_weights2=edge_weights+all_comb2_rank/max(all_comb2_rank)



E(graph_coex)$weight=edge_weights2
E(graph_coex)$width=log2(E(graph_coex)$weight)

```






```{r, transparentcolors}




new_colors=edge_color

for(i in 1:length(new_colors)){
percent=1/max(all_comb2_rank)*all_comb2_rank[i]

new_colors[i] <- adjustcolor(edge_color[i], alpha.f = percent)
}




```


```{r}

set.seed(1)
mod.copy_coex=cluster_fast_greedy(graph_coex, weights = E(graph_coex)$weight)
# set.seed(1)

# mod.copy_coex3=cluster_edge_betweenness(graph_coex, weights = E(graph_coex)$weight) 
# set.seed(1)
# 
# mod.copy_coex4=cluster_leading_eigen(graph_coex, weights = E(graph_coex)$weight) 
# At arpack.c:776 :ARPACK solver failed to converge (1001 iterations, 0/1 eigenvectors converged)Error in cluster_leading_eigen(graph_coex, weights = E(graph_coex)$weight) : 
#   At arpack.c:944 : ARPACK error, Maximum number of iterations reached
# set.seed(1)
# 
# mod.copy_coex4=cluster_spinglass(graph_coex, weights = E(graph_coex)$weight) 
# set.seed(1)
# 
# mod.copy_coex5=cluster_louvain(graph_coex, weights = E(graph_coex)$weight) 
# set.seed(1)
# mod.copy_coex6=cluster_label_prop(graph_coex, weights = E(graph_coex)$weight) 


lista.mod.copy=list()

lista.mod.copy$`1`=mod.copy_coex[[1]]
lista.mod.copy$`2`=mod.copy_coex[[2]]
lista.mod.copy$`3`=mod.copy_coex[[3]]
lista.mod.copy$`4`=mod.copy_coex[[4]]
lista.mod.copy$all=V(graph_coex)$name



weights <- ifelse(igraph::crossing(mod.copy_coex, graph_coex), 1, .5)
layout <- layout_with_kk(graph_coex, weights=weights)
```

```{r}



whitered=colorRampPalette(c("white", "red"))
bluewhite=colorRampPalette(c("blue", "white"))



cutcolors_pos=whitered(100)[as.numeric(cut(interesting_genes$median_log2FC[which(interesting_genes$median_log2FC>=0)],breaks = 100))]

cutcolors_neg=bluewhite(100)[as.numeric(cut(interesting_genes$median_log2FC[which(interesting_genes$median_log2FC<0)],breaks = 100))]


cutcolors_df_pos=data.frame("name"=interesting_genes$ensembl_gene_id[which(interesting_genes$median_log2FC>=0)],"colors"=cutcolors_pos)
cutcolors_df_neg=data.frame("name"=interesting_genes$ensembl_gene_id[which(interesting_genes$median_log2FC<0)],"colors"=cutcolors_neg)


rownames(cutcolors_df_neg)=cutcolors_df_neg$name
rownames(cutcolors_df_pos)=cutcolors_df_pos$name

cutcolors_df=rbind(cutcolors_df_neg,cutcolors_df_pos)
cutcolors_df=cutcolors_df[V(graph_coex)$name,]
```



```{r, fig.width=7, fig.height=5}
# pdf(width=7, height=5, file="network_transparent.pdf")
par(mar=c(0,0,0,4))
set.seed(1)
plot(graph_coex, layout=layout, vertex.label=nomi_vertex, vertex.color=as.character(cutcolors_df$colors), edge.color=new_colors, vertex.label.color=nomi_col)
colkey(rev(viridis(12)), clim=c(2,13), add=T, clab="length co-DE cluster", width=.7, length=.3, shift = .2, dist=(-1)*0.05)
colkey(c(bluewhite(100), whitered(100)), add=T, clab="median log2FC", width=.7, length=.3, clim=c(min(interesting_genes$median_log2FC),max(interesting_genes$median_log2FC)),  shift=(-1)*0.2, dist=(-1)*0.05 )

legend("topleft", 
       c("other chrs", "chr21"), bty="n",cex=1,
       pch=c(1,0), title="deregulated in:")

legend("bottomleft", 
       c("low", "high"), bty="n",pt.cex=c(1,3),
       pch=1, title="log2(occurences):",cex=1)







```



### Analysis of STRINGdb



```{r addingstring}
library(STRINGdb)

edges_matrix=as.data.frame(cbind(edges_names,edge_weights))
rownames(edges_matrix)=edges_matrix[,1]

string_db <- STRINGdb$new(version="11",species=9606,score_threshold=0, input_directory="" )

list_ds=as.data.frame(interesting_genes$ensembl_gene_id)
colnames(list_ds)="list_ds"

mapped <- string_db$map( list_ds, "list_ds", removeUnmappedRows = FALSE )

interactions=string_db$get_interactions(mapped$STRING_id )

interactions_merged=merge(interactions, mapped,by.x="from", by.y="STRING_id", all.x = TRUE)

colnames(interactions_merged)[4]="from_ens"
interactions_merged=merge(interactions_merged, mapped,by.x="to", by.y="STRING_id", all.x = TRUE)


graph_string=graph_from_data_frame(interactions_merged[,4:5], directed = FALSE, vertices = NULL)


firstsecond_string=paste(interactions_merged$from_ens, interactions_merged$list_ds, sep=".")

secondfirst_string=paste(interactions_merged$list_ds, interactions_merged$from_ens, sep=".")

combined_scores=rep(0, length(E(graph_coex)$weight))

names(combined_scores)=edges_names

combined_scores=as.data.frame(combined_scores)

combined_scores$int=rownames(combined_scores)

interactions_merged$firstsecond=firstsecond_string
interactions_merged$second_first=secondfirst_string

combined_scores_merged=merge(combined_scores, interactions_merged, by.x="int", by.y="firstsecond", all.x=TRUE)

combined_scores_merged=combined_scores_merged[,c(1, 5)]
colnames(combined_scores_merged)=c("interactions", "score1")

combined_scores_merged=merge(combined_scores_merged, interactions_merged, by.x="interactions", by.y="second_first", all.x=TRUE)

combined_scores_merged=combined_scores_merged[,c(1,2, 5)]
combined_scores_merged=combined_scores_merged[!duplicated(combined_scores_merged$interactions),]
rownames(combined_scores_merged)=combined_scores_merged$interactions

combined_scores_merged=combined_scores_merged[,c(2,3)]

combined_scores_merged$string_score=rowSums(combined_scores_merged, na.rm=T)


combined_scores_merged=combined_scores_merged[edges_names,]

# plot(as.numeric(E(graph_coex)$weight), combined_scores_merged$string_score, xlab="number of co-expression occurences", ylab="STRINGdb score")

combined_scores_merged$string_score=combined_scores_merged$string_score/1000

total_score=as.numeric(E(graph_coex)$weight)+ as.numeric(E(graph_coex)$weight)*combined_scores_merged$string_score



```


```{r}


string_interaction=interactions_merged[,4:5]




firstsecond_string=paste(interactions_merged$from_ens, interactions_merged$list_ds, sep=".")

secondfirst_string=paste(interactions_merged$list_ds, interactions_merged$from_ens, sep=".")

int_coex=paste(mat_coex[,1], mat_coex[,2], sep=".")


combined_scores_merged2=combined_scores_merged[c(intersect(firstsecond_string, int_coex), intersect(secondfirst_string, int_coex)),]

edges_matrix2=edges_matrix[rownames(combined_scores_merged2),]


# plot(jitter(as.numeric(as.character(edges_matrix2$edge_weights))), combined_scores_merged2$string_score)
# mean(as.numeric(as.character(edges_matrix[setdiff(edges_names, rownames(combined_scores_merged2)),2])))

# mean(as.numeric(as.character(edges_matrix2[which(combined_scores_merged2$string_score<0.4 & combined_scores_merged$string_score>0),2])), na.rm=T)
# mean(as.numeric(as.character(edges_matrix2[which(combined_scores_merged$string_score>=0.4),2])), na.rm=T)




```


We detected `r length(firstsecond_string)` interactions in StringDB between our DS genes, of whose `r length(intersect(firstsecond_string, int_coex))+length(intersect(secondfirst_string, int_coex))` were also found as consistently DE (`r round((length(intersect(firstsecond_string, int_coex))+length(intersect(secondfirst_string, int_coex)))/length(firstsecond_string)*100,0)` %).

### Overlap with GOs

```{r addingstring2}
library(STRINGdb)
list_coex2=lapply(names(all_comb2_rank)[which(all_comb2_rank>3)], function(x)(strsplit(x, split=".", fixed=T)[[1]]))
mat_coex2=do.call(rbind, list_coex2)



edges_names2=paste(mat_coex2[,1], mat_coex2[,2], sep=".")


edge_weights2=edge_weights[which(all_comb2_rank>3)]

edges_matrix2=as.data.frame(cbind(edges_names2,edge_weights2))
rownames(edges_matrix2)=edges_matrix2[,1]

string_db <- STRINGdb$new(version="11",species=9606,score_threshold=0, input_directory="" )

list_ds2=as.data.frame(all_comb2_rank_top_genes)
colnames(list_ds2)="list_ds"

mapped2 <- string_db$map( list_ds2, "list_ds", removeUnmappedRows = FALSE )

interactions2=string_db$get_interactions(mapped2$STRING_id )

interactions_merged2=merge(interactions2, mapped2,by.x="from", by.y="STRING_id", all.x = TRUE)

colnames(interactions_merged2)[4]="from_ens"
interactions_merged2=merge(interactions_merged2, mapped2,by.x="to", by.y="STRING_id", all.x = TRUE)




firstsecond_string2=paste(interactions_merged2$from_ens, interactions_merged2$list_ds, sep=".")

secondfirst_string2=paste(interactions_merged2$list_ds, interactions_merged2$from_ens, sep=".")

combined_scores2=rep(0, length(edges_names2))

names(combined_scores2)=edges_names2

combined_scores2=as.data.frame(combined_scores2)

combined_scores2$int=rownames(combined_scores2)

interactions_merged2$firstsecond=firstsecond_string2
interactions_merged2$second_first=secondfirst_string2

combined_scores_merged2=merge(combined_scores2, interactions_merged2, by.x="int", by.y="firstsecond", all.x=TRUE)

combined_scores_merged2=combined_scores_merged2[,c(1, 5)]
colnames(combined_scores_merged2)=c("interactions", "score1")

combined_scores_merged2=merge(combined_scores_merged2, interactions_merged2, by.x="interactions", by.y="second_first", all.x=TRUE)

combined_scores_merged2=combined_scores_merged2[,c(1,2, 5)]
combined_scores_merged2=combined_scores_merged2[!duplicated(combined_scores_merged2$interactions),]
rownames(combined_scores_merged2)=combined_scores_merged2$interactions
combined_scores_merged2=combined_scores_merged2[,c(2,3)]

combined_scores_merged2$string_score=rowSums(combined_scores_merged2, na.rm=T)


combined_scores_merged2=combined_scores_merged2[edges_names2,]

# plot(as.numeric(E(graph_coex)$weight), combined_scores_merged$string_score, xlab="number of co-expression occurences", ylab="STRINGdb score")

combined_scores_merged2$string_score=combined_scores_merged2$string_score/1000

# total_score=as.numeric(E(graph_coex)$weight)+ as.numeric(E(graph_coex)$weight)*combined_scores_merged$string_score



```


```{r}

interactions_graph=names(all_comb2_rank)[which(all_comb2_rank>3)]


# length(intersect(interactions_merged$firstsecond, interactions_graph))
# 
# length(intersect(interactions_merged$second_first, interactions_graph))
# 


indexes_common2=sort(unique(c(which(interactions_merged2$firstsecond %in% interactions_graph), which(interactions_merged2$second_first %in% interactions_graph))))

scores_db2=interactions_merged2$combined_score[indexes_common2]


interactions_merged2$selected=rep(NA, length(interactions_merged2$firstsecond))

interactions_merged2$selected[which(interactions_merged2$firstsecond %in% interactions_graph)]=interactions_merged2$firstsecond[which(interactions_merged2$firstsecond %in% interactions_graph)]

interactions_merged2$selected[which(interactions_merged2$second_first %in% interactions_graph)]=interactions_merged2$second_first[which(interactions_merged2$second_first %in% interactions_graph)]

names_db2=interactions_merged2$selected[indexes_common2]


names(scores_db2)=names_db2

# plot(as.numeric(all_comb2_rank[names_db2]), scores_db2)
# 
# length(unique(c(interactions_merged2$from_ens, interactions_merged2$list_ds)))

```


```{r}


go_df=getBM(attributes=c("ensembl_gene_id", "go_id"),filters="ensembl_gene_id", values=all_comb2_rank_top_genes, mart = human, useCache = F)
go_df=go_df[which(go_df$go_id!=""),]
go_df_split=split(go_df$go_id, go_df$ensembl_gene_id)
overlap_go=rep(NA, length(all_comb2_rank_top))
names(overlap_go)=names(all_comb2_rank_top)

for(i in 1:length(overlap_go)){
  overlap_go[i]=length(intersect(go_df_split[[mat_coex2[i,1]]], go_df_split[[mat_coex2[i,2]]]))
  
}
withgo=length(which(overlap_go>0))
  set.seed(1)
  nperm=1000
  
  overlaps_perms=rep(NA, nperm)
  for(j in 1:nperm){
    random_genes=sample(unique(unlist(list_array$names)), length(all_comb2_rank_top_genes))
    go_df_rand=getBM(attributes=c("ensembl_gene_id", "go_id"),filters="ensembl_gene_id", values=random_genes, mart = human, useCache = FALSE)
    go_df_rand=go_df_rand[which(go_df_rand$go_id!=""),]
    go_df_split_rand=split(go_df_rand$go_id, go_df_rand$ensembl_gene_id)
    overlap_go_rand=rep(NA, length(all_comb2_rank_top))
    
  
    combinazioni=combn(random_genes,2)
    combinazioni=t(combinazioni)
    rand_index=sample(1:dim(combinazioni)[1], length(all_comb2_rank_top))
    
    combinazioni=combinazioni[rand_index,]
    
    names(overlap_go_rand)=paste(combinazioni[,1], combinazioni[,2], sep=".")
    
    
    
    for(i in 1:length(overlap_go_rand)){
    overlap_go_rand[i]=length(intersect(go_df_split_rand[[combinazioni[i,1]]], go_df_split_rand[[combinazioni[i,2]]]))
    }
    
    overlaps_perms[j]=length(which(overlap_go_rand>0))
   # print(j) 
  }





overlap_go_db=rep(NA, length(all_comb2_rank[names_db2]))


names(overlap_go_db)=names(all_comb2_rank[names_db2])


for(i in 1:length(overlap_go_db)){
  
  touse=strsplit(names_db2[i], split=".", fixed=T)[[1]]
  overlap_go_db[i]=length(intersect(go_df_split[[touse[1]]], go_df_split[[touse[2]]]))
  
}

  touse=strsplit(names_db2[i], split=".", fixed=T)[[1]]


tot=unique(c(names(overlap_go[which(overlap_go>0)]),names_db2))


# length(unique(unlist(lapply(tot, FUN=function(x)(strsplit(x, split=".", fixed=T)[[1]])))))


```

The number of coDe with a GO was `r withgo/mean(overlaps_perms)` than expected by chance.


### Edge width proportional to string score (only shown edge with co-DE in at least 4 datsets)


```{r}


widths2=(combined_scores_merged$string_score)*3
#edge width proportional to stringscore
par(mar=c(0,0,0,4))
set.seed(1)
plot(graph_coex, layout=layout, vertex.label="", vertex.color=as.character(cutcolors_df$colors), edge.color=new_colors, edge.width=widths2)
legend("bottomright", 
       c("high", "low"), bty="n",lwd=c(3,1),
       , cex=1, title="stringDB score:",pch="_")
```


### Manhattan plot


```{r}
gene_frequency=sort(table(unlist(list_array.05_fc1.5_max500$names)), decreasing=T)

allgenes=as.data.frame(gene_frequency)

colnames(allgenes)=c("names", "occurences")


allgenes_anno=getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "chromosome_name", "band"),filters="ensembl_gene_id", values=allgenes$names, mart = human)


allgenes_anno$description=unlist(lapply(allgenes_anno$description, FUN=function(x)strsplit(x, split='[', fixed=T)[[1]][1]))


allgenes=merge(allgenes, allgenes_anno, all.x=TRUE, by.x="names", by.y="ensembl_gene_id")

allgenes=allgenes[!duplicated(allgenes$names),]

allgenes$median_log2FC=rep(NA, dim(allgenes)[1])
allgenes$sd_log2FC=rep(NA, dim(allgenes)[1])
allgenes$mean_log2FC=rep(NA, dim(allgenes)[1])


for(i in 1:dim(allgenes)[1]){
  ensgene=allgenes$names[i]
  fcs=find_fold_changes(ensgene, list_array.05_fc1.5_max500)
  allgenes$median_log2FC[i]=median(fcs)
  allgenes$sd_log2FC[i]=sd(fcs)
  allgenes$mean_log2FC[i]=mean(fcs)
  
}

allgenes$pseudotscore=allgenes$mean_log2FC/(allgenes$sd_log2FC/sqrt(allgenes$occurences))

rownames(allgenes)=allgenes$names
allgenes$consistently_DE=rep("no", length(allgenes$names))

allgenes$consistently_DE[which(allgenes$names %in% rownames(interesting_genes))]="yes"
```


```{r, fig.width=17, fig.height=4.5}

allgenes_autosomes=allgenes[which(allgenes$chromosome_name %in% as.character(1:22)),]
allgenes_sexchr=allgenes[which(allgenes$chromosome_name %in% c("X", "Y")),]


allgenes_autosomes$chromosome_name=as.numeric(allgenes_autosomes$chromosome_name)
allgenes_autosomes=allgenes_autosomes[order(allgenes_autosomes$chromosome_name, allgenes_autosomes$band),]

allgenes_sexchr=allgenes_sexchr[order(allgenes_sexchr$chromosome_name, allgenes_sexchr$band),]


allgenes_ordered=rbind(allgenes_autosomes, allgenes_sexchr)

allgenes_ordered=allgenes_ordered[which(!is.na(allgenes_ordered$sd_log2FC)),]

transblue=colorRampPalette(c("transparent", "blue"))

colors_mahattan=transblue(max(ceiling(log2(allgenes_ordered$occurences))))[ceiling(log2(allgenes_ordered$occurences))]
colors_mahattan1=transblue(max(allgenes_ordered$occurences))[allgenes_ordered$occurences]

number_vec=allgenes_ordered$occurences
color="blue"
create_gradient=function(number_vec, color){
  library(scales)
  toreturn=rep(NA, length(number_vec))
  
  for(i in 1:length(toreturn)){
    toreturn[i]=alpha(color, number_vec[i]/max(number_vec))
  }
  return(toreturn)
}

colors_mahattan=create_gradient(allgenes_ordered$occurences, "blue")
colors_mahattan2=create_gradient(ceiling(log2(allgenes_ordered$occurences)), "blue")

plot(1:length(allgenes_ordered$median_log2FC), allgenes_ordered$median_log2FC, pch=16, ylab=expression(paste("log"[2], "FC (median)")), xlab="HSA chromosome", xaxt="n", type = "n")
# 
# 
points(which(allgenes_ordered$consistently_DE =="no"), allgenes_ordered$median_log2FC[which(allgenes_ordered$consistently_DE =="no")], pch=16, cex=.5, col=colors_mahattan2[which(allgenes_ordered$consistently_DE =="no")])
points(which(allgenes_ordered$consistently_DE =="yes"), allgenes_ordered$median_log2FC[which(allgenes_ordered$consistently_DE =="yes")], pch=16, cex=1,col=colors_mahattan2[which(allgenes_ordered$consistently_DE =="no")])

axis_lab=rep(NA,length(allgenes_ordered$median_log2FC))

for(i in 1:length(unique(allgenes_ordered$chromosome_name))){
  axis_lab[round(mean(which(allgenes_ordered$chromosome_name %in% unique(allgenes_ordered$chromosome_name)[i])))]=unique(allgenes_ordered$chromosome_name)[i]
}


axis(1, which(!is.na(axis_lab)), labels=axis_lab[which(!is.na(axis_lab))])
abline(h=0)
abline(h=log2(1.5), lty="dashed", col="gray")
abline(h=-log2(1.5), lty="dashed", col="gray")

for(i in unique(allgenes_ordered$chromosome_name)){
  abline(v=(which(allgenes_ordered$chromosome_name %in% i)[1]-0.5))
}



```

```{r, fig.width=17, fig.height=4.5}
bluetrans=colorRampPalette(c("blue", "transparent"))

allgenes_ordered=allgenes_ordered[which(allgenes_ordered$sd_log2FC>0),]
cutcolors_pos_pts=transblue(max(rank(allgenes_ordered$pseudotscore[which(allgenes_ordered$pseudotscore>=0)])))[rank(allgenes_ordered$pseudotscore[which(allgenes_ordered$pseudotscore>=0)])]
                                                                                                               
                                                                                                               
                                                                                                               
cutcolors_neg_pts=bluetrans(max(rank(allgenes_ordered$pseudotscore[which(allgenes_ordered$pseudotscore<0)])))[rank(allgenes_ordered$pseudotscore[which(allgenes_ordered$pseudotscore<0)])]


cutcolors_df_pos_pts=data.frame("name"=allgenes_ordered$names[which(allgenes_ordered$pseudotscore>=0)],"colors"=cutcolors_pos_pts)
cutcolors_df_neg_pts=data.frame("name"=allgenes_ordered$names[which(allgenes_ordered$pseudotscore<0)],"colors"=cutcolors_neg_pts)


rownames(cutcolors_df_neg_pts)=cutcolors_df_neg_pts$name
rownames(cutcolors_df_pos_pts)=cutcolors_df_pos_pts$name

cutcolors_df_pts=rbind(cutcolors_df_neg_pts,cutcolors_df_pos_pts)

cutcolors_df_pts=cutcolors_df_pts[as.character(allgenes_ordered$names),]

cexes=allgenes_ordered$consistently_DE
cexes[which(cexes=="no")]=0.5
cexes[which(cexes=="yes")]=1
cexes=as.numeric(cexes)
plot(1:length(allgenes_ordered$median_log2FC), allgenes_ordered$median_log2FC, pch=16, ylab=expression(paste("log"[2], "FC (median)")), xlab="HSA chromosome", xaxt="n",col=as.character(cutcolors_df_pts$colors), cex=cexes)


axis_lab=rep(NA,length(allgenes_ordered$median_log2FC))

for(i in 1:length(unique(allgenes_ordered$chromosome_name))){
  axis_lab[round(mean(which(allgenes_ordered$chromosome_name %in% unique(allgenes_ordered$chromosome_name)[i])))]=unique(allgenes_ordered$chromosome_name)[i]
}


axis(1, which(!is.na(axis_lab)), labels=axis_lab[which(!is.na(axis_lab))])
abline(h=0)
abline(h=log2(1.5), lty="dashed", col="gray")
abline(h=-log2(1.5), lty="dashed", col="gray")

for(i in unique(allgenes_ordered$chromosome_name)){
  abline(v=(which(allgenes_ordered$chromosome_name %in% i)[1]-0.5))
}

```


`r length(which(interesting_genes$mean_log2FC>0))` genes out of `r dim(interesting_genes)[1]` were on average upregulated.



### Distribution of pseudo-t-score

```{r}
interesting_genes$pseudotscore=interesting_genes$mean_log2FC/(interesting_genes$sd_log2FC/sqrt(interesting_genes$occurences))

h=hist(interesting_genes$pseudotscore, breaks = 200, xlab="pseudo t-score", main="", plot=F)


cuts <- as.numeric(cut(h$breaks, c(-Inf,-1.96,1.96,Inf)))

colori=rep("black", length(cuts))

colori[which(cuts==2)]="white"
plot(h, col=colori, breaks = 200, xlab="pseudo t-score", main="")
```





### Graph colored with standardized pseudo t-score


```{r}
whitered=colorRampPalette(c("white", "red"))
bluewhite=colorRampPalette(c("blue", "white"))


cutcolors_pos=whitered(100)[as.numeric(cut(interesting_genes$pseudotscore[which(interesting_genes$mean_log2FC>=0)],breaks = 100))]

cutcolors_neg=bluewhite(100)[as.numeric(cut(interesting_genes$pseudotscore[which(interesting_genes$mean_log2FC<0)],breaks = 100))]


cutcolors_df_pos=data.frame("name"=interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC>=0)],"colors"=cutcolors_pos)
cutcolors_df_neg=data.frame("name"=interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC<0)],"colors"=cutcolors_neg)


rownames(cutcolors_df_neg)=cutcolors_df_neg$name
rownames(cutcolors_df_pos)=cutcolors_df_pos$name

cutcolors_df=rbind(cutcolors_df_neg,cutcolors_df_pos)
cutcolors_df=cutcolors_df[V(graph_coex)$name,]
```



```{r, fig.height=14, fig.width=14}


par(mar=c(0,0,0,5))
set.seed(1)
plot(graph_coex, layout=layout, vertex.label=nomi_vertex, vertex.color=as.character(cutcolors_df$colors), edge.color=new_colors, vertex.label.color=nomi_col)
colkey(rev(viridis(12)), clim=c(2,13), add=T, clab="length co-DE cluster", width=.7, length=.3, shift = .2, dist=(-1)*0.05)
colkey(c(bluewhite(100), whitered(100)), add=T, clab="standardized pseudo t-score", width=.7, length=.3, clim=c(min(interesting_genes$pseudotscore),max(interesting_genes$pseudotscore)),  shift=(-1)*0.2, dist=(-1)*0.05 )

legend("topleft", 
       c("other chrs", "chr21"), bty="n",cex=1,
       pch=c(1,0), title="deregulated in:")

legend("bottomleft", 
       c("low", "high"), bty="n",pt.cex=c(1,3),
       pch=1, title="log2(occurences):",cex=1)












highest_changing_genes_table=interesting_genes[which(abs(interesting_genes$pseudotscore)>1.96),]


highest_changing_genes=highest_changing_genes_table$ensembl_gene_id

#highest_changing_genes_table[intersect(rownames(interesting_genes),highest_changing_genes),]


```

We detected `r length(highest_changing_genes)` with a significant pseudo t-score.


```{r highest}
tab_int=interesting_genes[highest_changing_genes,]
tab_int=tab_int[order(tab_int$pseudotscore, decreasing = T),]

options(knitr.table.format = "html") 
kable(tab_int, "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) %>%
  scroll_box(width = "500px", height = "200px")

```



### Enrichment analysis for genes with high pseudo-t-score

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
lista_ensembl=highest_changing_genes
lista.id <- bitr(lista_ensembl, fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)$ENTREZID


```


#### GO

```{r}
ego <- enrichGO(gene          = lista.id,
                OrgDb         = org.Hs.eg.db,
                ont           = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
        readable      = TRUE)
```


```{r}
dotplot(ego, showCategory=100)
IFNgenes=c("IFNAR2","IFNAR1","RRP1B","EIF2AK2","IRF7","IFIT3","ILF3","BST2","TRIM5","RTP4","IFI27","IFIT1B","IRGM", "USP18","UBE2G2")
neutrophil=c("PFKL","CCT8","CSTB","PDXK","CAT","METTL7A","PLAC8","ITGB2","CD55","VCL","ASAH1","BST2","S100A9","CD47")
```

### Interactor level

```{r}
library(STRINGdb)
string_db2 <- STRINGdb$new(version="11",species=9606,score_threshold=900, input_directory="" )


list_ds=as.data.frame(all_genes$ensembl_gene_id[grep("21", all_genes$chromosome_name)])
colnames(list_ds)="list_ds"



mapped <- string_db2$map( list_ds, "list_ds", removeUnmappedRows = FALSE )


neighbors=c()
for (i in 1:length(mapped$STRING_id[!(is.na(mapped$STRING_id))])) {
  tryCatch({
    neighbors=append(neighbors, string_db2$get_neighbors(mapped$STRING_id[!(is.na(mapped$STRING_id))][i] ))
  }, error=function(e){})
}

neighbors=unique(neighbors)

neighbors_ens=unique(getBM(attributes=c("ensembl_peptide_id", "ensembl_gene_id"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors, FUN=function(x)(substring(x,6)))), mart = human)$ensembl_gene_id)



list_ds2=as.data.frame(neighbors)
colnames(list_ds2)="list_ds2"



mapped2 <- string_db2$map( list_ds2, "list_ds2", removeUnmappedRows = FALSE )


neighbors2=c()
for (i in 1:length(mapped2$STRING_id[!(is.na(mapped2$STRING_id))])) {
  tryCatch({
    neighbors2=append(neighbors2, string_db2$get_neighbors(mapped2$STRING_id[!(is.na(mapped2$STRING_id))][i] ))
  }, error=function(e){})
}

neighbors2=unique(neighbors2)

neighbors_ens2=unique(getBM(attributes=c("ensembl_peptide_id", "ensembl_gene_id"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors2, FUN=function(x)(substring(x,6)))), mart = human)$ensembl_gene_id)


###


list_ds3=as.data.frame(neighbors2)
colnames(list_ds3)="list_ds3"



mapped3 <- string_db2$map( list_ds3, "list_ds3", removeUnmappedRows = FALSE )


neighbors3=c()
for (i in 1:length(mapped3$STRING_id[!(is.na(mapped3$STRING_id))])) {
  tryCatch({
    neighbors3=append(neighbors3, string_db2$get_neighbors(mapped3$STRING_id[!(is.na(mapped3$STRING_id))][i] ))
  }, error=function(e){})
}

neighbors3=unique(neighbors3)

neighbors_ens3=unique(getBM(attributes=c("ensembl_peptide_id", "ensembl_gene_id"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors3, FUN=function(x)(substring(x,6)))), mart = human)$ensembl_gene_id)



###


list_ds4=as.data.frame(neighbors3)
colnames(list_ds4)="list_ds4"



mapped4 <- string_db2$map( list_ds4, "list_ds4", removeUnmappedRows = FALSE )


neighbors4=c()
for (i in 1:length(mapped4$STRING_id[!(is.na(mapped4$STRING_id))])) {
  tryCatch({
    neighbors4=append(neighbors4, string_db2$get_neighbors(mapped4$STRING_id[!(is.na(mapped4$STRING_id))][i] ))
  }, error=function(e){})
}

neighbors4=unique(neighbors4)

neighbors_ens4=unique(getBM(attributes=c("ensembl_peptide_id", "ensembl_gene_id"),filters="ensembl_peptide_id", values=unlist(lapply(neighbors4, FUN=function(x)(substring(x,6)))), mart = human)$ensembl_gene_id)

interesting_genes$interactor_level=rep("5+", dim(interesting_genes)[1])

interesting_genes$interactor_level[which(interesting_genes$ensembl_gene_id %in% intersect(interesting_genes$ensembl_gene_id, neighbors_ens4))]=4

interesting_genes$interactor_level[which(interesting_genes$ensembl_gene_id %in% intersect(interesting_genes$ensembl_gene_id, neighbors_ens3))]=3

interesting_genes$interactor_level[which(interesting_genes$ensembl_gene_id %in% intersect(interesting_genes$ensembl_gene_id, neighbors_ens2))]=2

interesting_genes$interactor_level[which(interesting_genes$ensembl_gene_id %in% intersect(interesting_genes$ensembl_gene_id, neighbors_ens))]=1

interesting_genes$interactor_level[which(interesting_genes$chromosome_name=="21")]=0

interesting_splitted=split(interesting_genes, interesting_genes$interactor_level)

occurrences_interactorlevel=rep(NA, length(interesting_splitted))
for(i in 1:length(interesting_splitted)){
  occurrences_interactorlevel[i]=mean(interesting_splitted[[i]]$occurences)
}

# > occurrences_interactorlevel
# [1] 7.813725 4.640777 4.581699 4.702703 4.500000 4.435644


occurrences_pseudotscore=rep(NA, length(interesting_splitted))
for(i in 1:length(interesting_splitted)){
  occurrences_pseudotscore[i]=mean(interesting_splitted[[i]]$pseudotscore)
}
# > occurrences_pseudotscore
# [1]  5.6213806155  3.0474310263  0.8454890448  1.5996111159  1.3980162160 -0.0006106038


occurrences_mean_log2FC=rep(NA, length(interesting_splitted))
for(i in 1:length(interesting_splitted)){
  occurrences_mean_log2FC[i]=mean(interesting_splitted[[i]]$mean_log2FC)
}
# > occurrences_mean_log2FC
# [1] 0.70914303 0.24463838 0.24968291 0.25619759 0.16968330 0.03433787

occurrences_median_log2FC=rep(NA, length(interesting_splitted))
for(i in 1:length(interesting_splitted)){
  occurrences_median_log2FC[i]=mean(interesting_splitted[[i]]$median_log2FC)
}


# ggplot(interesting_genes, aes(y=pseudotscore, x=as.factor(interactor_level))) +xlab("interactor level")  +ylab("pseudo t-score")+geom_bar(fill="steelblue", stat = "summary", fun.y = "mean")+
#   theme_minimal() +geom_errorbar(aes(ymin=pseudotscore-sd, ymax=pseudotscore+sd), width=.2,
#                                     position=position_dodge(.9)) 
# 
# 
# 


occurrences_mean_abslog2FC=rep(NA, length(interesting_splitted))
for(i in 1:length(interesting_splitted)){
  occurrences_mean_abslog2FC[i]=mean(abs(interesting_splitted[[i]]$mean_log2FC))
}



#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
# for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
#to be summariezed
# groupnames : vector of column names to be used as
# grouping variables
data_summary <- function(data, varname, groupnames){
  library(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      se = sd(x[[col]], na.rm=TRUE)/sqrt(length(x[[col]])))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}


df2 <- data_summary(interesting_genes, varname="pseudotscore", 
                     groupnames="interactor_level")
# Convert dose to a factor variable
df2$dose=as.factor(df2$interactor_level)



p<- ggplot(df2, aes(x=interactor_level, y=pseudotscore, fill=interactor_level)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=pseudotscore, ymax=pseudotscore+se), width=.2,
                position=position_dodge(.9)) +theme_minimal()
print(p)
```

### co-DE and TADs

```{r}
tadsboundaries_cortex=read.csv("cortex_boundaries_dixon.csv", header=T)

tadsboundaries_hESCs=read.csv("hESC_tads_dixon.csv", header=F)

```

```{r, fig.width=10}
# human = useMart("ensembl", dataset = "hsapiens_gene_ensembl",host="oct2018.archive.ensembl.org")
human = useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl",
                   mirror = "useast")


ensembl = useMart("ENSEMBL_MART_ENSEMBL",
   dataset="hsapiens_gene_ensembl",
   host="may2009.archive.ensembl.org",
   path="/biomart/martservice",
   archive=FALSE)


genes_coords=getBM(attributes=c("ensembl_gene_id",  "chromosome_name", "start_position", "end_position"),filters="ensembl_gene_id", values=interesting_genes$ensembl_gene_id, mart = ensembl)

genes_coords1=getBM(attributes=c("ensembl_gene_id",  "chromosome_name", "start_position", "end_position"),filters="ensembl_gene_id", values=interesting_genes$ensembl_gene_id, mart = human)

# hg18conv=getLDS(attributes = c("start_position"), filters = "start_position", values = genes_coords1$start_position, mart = human, attributesL = c("start_position"), martL = ensembl)


rownames(genes_coords)=genes_coords$ensembl_gene_id
tadsboundaries_hESCs$name=rownames(tadsboundaries_hESCs)
tad_object=tadsboundaries_hESCs
coord_object=genes_coords
# gene="ENSG00000142168"
find_tad_number=function(gene, tad_object, coord_object){
  
  subset=coord_object[gene,]
  chr=paste0("chr", subset$chromosome_name)
  subset_tad=tad_object[which(tad_object$V1==chr),]
  
    subset_tad=subset_tad[which(subset$start_position >=  subset_tad$V2),]
        subset_tad=subset_tad[which(subset$start_position <=  subset_tad$V3),]
    if(dim(subset_tad)[1]>1){subset_tad=subset_tad[which(subset$end_position <=  subset_tad$V3),]}
   #what matters is where the gene starts, only if there is more than one results look at the end position

    return(subset_tad$name)
}


tad_assignment=unlist(sapply(rownames(coord_object), FUN=function(x)(find_tad_number(x, tad_object, coord_object))))

tad_assignment=tad_assignment[rownames(interesting_genes)]

interesting_genes$tad=as.numeric(tad_assignment)

chrs=c(as.character(1:22), "X")
tads_per_chr=c()
for(i in chrs){
  tads_per_chr[i]=length(sort(table(interesting_genes[which(interesting_genes$chromosome_name==i),"tad"])))
}

# NCBI Build 36/hg18
# 
# listMarts(host="may2009.archive.ensembl.org", path="/biomart/martservice",archive=FALSE)
# 
# listMarts(host="may2009.archive.ensembl.org",
#            path="/biomart/martservice",archive=FALSE)


chrsizes=read.table("chrsizehg18.txt", sep="\t")


tadsperchr=table(tad_object$V1)

names(tadsperchr)=sapply(names(tadsperchr), FUN=function(x)(substr(x, start=4, stop=nchar(x))))

tadsperchr=tadsperchr[chrs]

barplot(rbind(tads_per_chr,tadsperchr), ylab="number of TADs")
legend("topright",c("all TADs", "TADs with DE gene(s)"), fill=c("lightgray", "gray27"), bty="n")



for(i in chrs){
  length(sort(table(interesting_genes[which(interesting_genes$chromosome_name==i),"tad"])))
}


frequency_tads=as.data.frame(table(tad_assignment))

merged=merge(frequency_tads,tad_object, by.x="tad_assignment", by.y="name", all.x=TRUE)

genes_tad_chr4=merged[,c(2,3)]





p <- ggplot(genes_tad_chr4, aes(x=V1, y=Freq)) + geom_boxplot()+ geom_jitter(shape=16, position=position_jitter(0.2))+ylab("DE genes per TAD")+ xlab("")  + theme_bw()
# Box plot with jittered points
# 0.2 : degree of jitter in x direction
p
```


```{r}
tadsmorethan1=names(table(tad_assignment)[which(table(tad_assignment)>1)])

interesting_genescotad=interesting_genes[which(interesting_genes$tad %in% tadsmorethan1),]


interesting_genescotad=interesting_genescotad[, c(1, dim(interesting_genescotad)[2])]

interesting_genescotad=split(interesting_genescotad$ensembl_gene_id, interesting_genescotad$tad)


combinazioni_tad=lapply(interesting_genescotad, FUN = function(x)(if(length(x)>1)(t(combn(x,2)))))



dimensioni_tad=unlist(lapply(combinazioni_tad, FUN=function(x) dim(x)[1]))
int_matrix_tad=as.data.frame(matrix(rep(NA, 3*sum(dimensioni_tad)), ncol=3))
contatore=1
for(i in 1:length(dimensioni_tad)){
  int_matrix_tad[contatore:(contatore+dimensioni_tad[i]-1),c(1,2)]=combinazioni_tad[[i]]
  int_matrix_tad[contatore:(contatore+dimensioni_tad[i]-1),3]=names(dimensioni_tad)[i]
  contatore=contatore+dimensioni_tad[i]
}
int_matrix_nometadata_tad=int_matrix_tad[,c(1,2)]
#order couple of interaction alphabetically
int_matrix_nometadata_tad=as.list(as.data.frame(t(int_matrix_nometadata_tad)))
int_matrix_nometadata_tad=lapply(int_matrix_nometadata_tad, sort)
unlisted=unlist(int_matrix_nometadata_tad)
unlisted=as.character(unlisted)
matrix_int_nomet_tad=matrix(unlisted, ncol=2, byrow=T)
primo_tad=paste(matrix_int_nomet_tad[,1], matrix_int_nomet_tad[,2]  , sep=".")

# 
# 
# intersect(edges_matrix2$edges_names2, primo_tad)
# 
# 
# 
# 
# 

combinazioni_tad3=lapply(interesting_genescotad, FUN = function(x)(if(length(x)>2)(t(combn(x,3)))))
combinazioni_tad3=combinazioni_tad3[!(unlist(lapply(combinazioni_tad3, is.null)))]

dimensioni_tad3=unlist(lapply(combinazioni_tad3, FUN=function(x) dim(x)[1]))
int_matrix_tad3=as.data.frame(matrix(rep(NA, 4*sum(dimensioni_tad)), ncol=4))
contatore=1
for(i in 1:length(dimensioni_tad3)){
  int_matrix_tad3[contatore:(contatore+dimensioni_tad3[i]-1),c(1,2,3)]=combinazioni_tad3[[i]]
  int_matrix_tad3[contatore:(contatore+dimensioni_tad3[i]-1),4]=names(dimensioni_tad3)[i]
  contatore=contatore+dimensioni_tad3[i]
}
int_matrix_nometadata_tad3=int_matrix_tad3[,c(1,2,3)]
#order couple of interaction alphabetically
int_matrix_nometadata_tad3=as.list(as.data.frame(t(int_matrix_nometadata_tad3)))
int_matrix_nometadata_tad3=lapply(int_matrix_nometadata_tad3, sort)
unlisted=unlist(int_matrix_nometadata_tad3)
unlisted=as.character(unlisted)
matrix_int_nomet_tad3=matrix(unlisted, ncol=3, byrow=T)
primo_tad3=paste(matrix_int_nomet_tad3[,1], matrix_int_nomet_tad3[,2]  , matrix_int_nomet_tad3[,3]  ,sep=".")

# intersect(names(all_comb3_rank_top), primo_tad3)
# 
# 
# 

combinazioni_tad4=lapply(interesting_genescotad, FUN = function(x)(if(length(x)>3)(t(combn(x,4)))))
combinazioni_tad4=combinazioni_tad4[!(unlist(lapply(combinazioni_tad4, is.null)))]

dimensioni_tad4=unlist(lapply(combinazioni_tad4, FUN=function(x) dim(x)[1]))
int_matrix_tad4=as.data.frame(matrix(rep(NA, 5*sum(dimensioni_tad)), ncol=5))
contatore=1
for(i in 1:length(dimensioni_tad4)){
  int_matrix_tad4[contatore:(contatore+dimensioni_tad4[i]-1),c(1,2,3,4)]=combinazioni_tad4[[i]]
  int_matrix_tad4[contatore:(contatore+dimensioni_tad4[i]-1),5]=names(dimensioni_tad4)[i]
  contatore=contatore+dimensioni_tad4[i]
}
int_matrix_nometadata_tad4=int_matrix_tad4[,c(1,2,3,4)]
#order couple of interaction alphabetically
int_matrix_nometadata_tad4=as.list(as.data.frame(t(int_matrix_nometadata_tad4)))
int_matrix_nometadata_tad4=lapply(int_matrix_nometadata_tad4, sort)
unlisted=unlist(int_matrix_nometadata_tad4)
unlisted=as.character(unlisted)
matrix_int_nomet_tad4=matrix(unlisted, ncol=4, byrow=T)
primo_tad4=paste(matrix_int_nomet_tad4[,1], matrix_int_nomet_tad4[,2]  , matrix_int_nomet_tad4[,3]  ,matrix_int_nomet_tad4[,4]  ,sep=".")
# intersect(names(all_comb4_rank_top), primo_tad4)
```

```{r}
lista=intersect(names(all_comb3_rank_top), primo_tad3)
intriplets=unique(unlist(lapply(lista, FUN=function(x)(strsplit(x, ".", fixed=TRUE)[[1]]))))

lista2=intersect(names(all_comb2_rank_top), primo_tad)

induplet=unique(unlist(lapply(lista2, FUN=function(x)(strsplit(x, ".", fixed=TRUE)[[1]]))))

duplet_matrix=interesting_genes[induplet,]

interactionwithstringdb=rep(0, length(edges_matrix2$edges_names2))

interactionwithstringdb[which(edges_matrix2$edges_names2 %in% firstsecond_string)]=1

interactionwithstringdb[which(edges_matrix2$edges_names2 %in% secondfirst_string)]=1



list_ds=as.data.frame(interesting_genes$ensembl_gene_id)
colnames(list_ds)="list_ds"

mapped <- string_db$map( list_ds, "list_ds", removeUnmappedRows = FALSE )

interactions=string_db$get_interactions(mapped$STRING_id )

interactions_merged=merge(interactions, mapped,by.x="from", by.y="STRING_id", all.x = TRUE)

colnames(interactions_merged)[4]="from_ens"
interactions_merged=merge(interactions_merged, mapped,by.x="to", by.y="STRING_id", all.x = TRUE)

consistent_interactions=interactions_merged[c(which(firstsecond_string %in% edges_matrix2$edges_names2), which(secondfirst_string %in% edges_matrix2$edges_names2)),]

```

`r dim(consistent_interactions)[1]` are both co-DE events both in STRINGdb and GO.

```{r}
# library(gplots)
# library(RColorBrewer)
# 
# heatmap.2(as.matrix(consistent_interactions[,c(3:5)]), scale="none", density.info="none", trace="none",srtCol=45, key.xlab ="score",margins=c(9,0),col=colorRampPalette(brewer.pal(9, "OrRd"))(256), labRow = "")
```

```{r}
interactionwithGO=rep(0, length(edges_matrix2$edges_names2))

interactionwithGO[which(edges_matrix2$edges_names2 %in% names(overlap_go[which(overlap_go>0)]))]=1

interactionwithTAD=rep(0, length(edges_matrix2$edges_names2))

interactionwithTAD[which(edges_matrix2$edges_names2 %in% primo_tad)]=1

table_edges=cbind(interactionwithstringdb,interactionwithGO,interactionwithTAD)
colnames(table_edges)=c("STRINGdb", "GO", "TAD")

```

```{r}
heatmap.2(table_edges, scale="none", density.info="none", trace="none",srtCol=45, key.xlab ="score",col=colorRampPalette(brewer.pal(9, "OrRd"))(256),  margins=c(8,4),key=FALSE, dendrogram="none", ylab="co-DE pairs", labRow = "", cexCol=1)
```


## Disease association



```{r}
library(disgenet2r)


genes=unique(interesting_genes$hgnc_symbol)

gene_disease <- gene2disease(
  gene     = genes,
  score =c(0, 1), database="ALL",
  verbose  = TRUE
)
results=extract(gene_disease)


write.csv(results, file="result_comorbidity_top500.csv")



disease_class=sort(table(results$disease_class_name), decreasing=T)

# write.csv(as.data.frame(disease_class), file="disease_class.csv")
# head(disease_class)          
          
disease_name=sort(table(results$disease_name), decreasing=T)

# write.csv(as.data.frame(disease_name), file="disease_name.csv")




```





```{r}
dpis=results[,c("gene_symbol", "geneid", "uniprotid", "gene_pli", "gene_dpi", "gene_dsi")]

dpis=dpis[!(duplicated(dpis$gene_symbol)),]
# list_diseases_sig=list_diseases[as.character(pvalues_01$Var1)]



comorb_merged=merge(interesting_genes, dpis, by.x="hgnc_symbol", by.y="gene_symbol")


comorb_merged=comorb_merged[order(comorb_merged$gene_dpi, decreasing = T),]

# plot(comorb_merged$median_log2FC,comorb_merged$gene_dpi, xlab="log2FC (median)", ylab="disease pleiotropic index", col=alpha("black", 0.3), pch=16)
# abline(v=log2(1.5), lty="dashed", col="red")
# abline(v=-log2(1.5), lty="dashed", col="red")
# 
# 
# plot(comorb_merged$median_log2FC,comorb_merged$gene_dpi, xlab="log2FC (median)", ylab="disease pleiotropic index", col=alpha("black", 0.3), pch=16)

# plot(comorb_merged$entropy_macro,comorb_merged$gene_dpi)


calc_mean_dpi=function(genes){
  gene_disease <- gene2disease(
  gene     = genes,
  score =c(0, 1), database="ALL",
  verbose  = FALSE
)
results=extract(gene_disease)
dpis=results[,c("gene_symbol", "geneid", "uniprotid", "gene_pli", "gene_dpi", "gene_dsi")]
dpis=dpis[!(duplicated(dpis$gene_symbol)),]

return(mean(dpis$gene_dpi, na.rm=T))
}
```


```{r}
# boxplot(comorb_merged$gene_dpi,
#         
#         comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% lista1_bm$hgnc_symbol)],
#         comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% lista2_bm$hgnc_symbol)],
# comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% lista3_bm$hgnc_symbol)])

nperm=1000
totest=lista1_bm$hgnc_symbol

meanall=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest)], na.rm=T)
permvalues=rep(NA, length(nperm))
set.seed(1)
for(i in 1:nperm){
  setsample=sample(comorb_merged$hgnc_symbol, length(lista2_bm$hgnc_symbol))
  permvalues[i]=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% setsample)], na.rm=T)
}
#(sum((permvalues > meanall))+1)/(nperm+1)



totest=lista3_bm$hgnc_symbol

meanall=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest)], na.rm=T)
permvalues=rep(NA, length(nperm))
set.seed(1)
for(i in 1:nperm){
  setsample=sample(comorb_merged$hgnc_symbol, length(lista2_bm$hgnc_symbol))
  permvalues[i]=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% setsample)], na.rm=T)
}
#(sum((permvalues > meanall))+1)/(nperm+1)
```

```{r}

genes=unique(interesting_genes$hgnc_symbol)

gene_disease <- gene2disease(
  gene     = genes,
  score =c(0, 1), database="ALL",
  verbose  = TRUE
)
results=extract(gene_disease)


gene_disease2 <- gene2disease(
  gene     = genes,
  score =c(0.4, 1), database="CURATED",
  verbose  = TRUE
)
results2=extract(gene_disease2)




gene_disease3 <- gene2disease(
  gene     = c(lista1_bm$hgnc_symbol, lista2_bm$hgnc_symbol,lista3_bm$hgnc_symbol),
  score =c(0.4, 1), database="CURATED",
  verbose  = TRUE
)
```



```{r}




nperm=200
totest=interesting_genes$hgnc_symbol

totest1=interesting_genes$hgnc_symbol[which(interesting_genes$chromosome_name=="21")]
totest2=setdiff(intersect(neighbors_hsa21, totest), totest1)

totest3=setdiff(interesting_genes$hgnc_symbol, c(totest1, totest2))
meanall=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest)], na.rm=T)
permvalues=rep(NA, length(nperm))
set.seed(1)
for(i in 1:nperm){
  setsample=sample(allgenes$hgnc_symbol, length(totest))
  permvalues[i]=calc_mean_dpi(setsample)
#  print(i)
}
(sum((permvalues > meanall))+1)/(nperm+1)


mean1=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest1)], na.rm=T)
mean2=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest2)], na.rm=T)
mean3=mean(comorb_merged$gene_dpi[which(comorb_merged$hgnc_symbol %in% totest3)], na.rm=T)


boxplot(list(meanall,permvalues), ylab="disease pleiotropic index (mean)", xaxt="n", col=c("blue", "red"))
axis(1, c(1,2),c("top 500 DE genes", "1000 permutations"))
```



```{r}
cluster1_GDAs=results2[results2$gene_symbol %in% lista1_bm$hgnc_symbol,]

kable(cluster1_GDAs[order(cluster1_GDAs$score, decreasing=T),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 



```


```{r}
cluster2_GDAs=results2[results2$gene_symbol %in% lista2_bm$hgnc_symbol,]

kable(cluster2_GDAs[order(cluster2_GDAs$score, decreasing=T),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 
```

```{r}
cluster3_GDAs=results2[results2$gene_symbol %in% lista3_bm$hgnc_symbol,]



kable(cluster3_GDAs[order(cluster3_GDAs$score, decreasing=T),], "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 

```


```{r}
plot(gene_disease3, class="Heatmap")
plot(gene_disease3, class="DiseaseClass")

```


### GO/DO similarity

```{r}
library(GOSemSim)

hsGO2 <- godata('org.Hs.eg.db', keytype = "ENSEMBL", ont="BP")
hsGO_mf <- godata('org.Hs.eg.db', keytype = "ENSEMBL", ont="MF")
hsGO_cc <- godata('org.Hs.eg.db', keytype = "ENSEMBL", ont="CC")
gene_sims_mf=mgeneSim(rownames(interesting_genes), semData=hsGO_mf, measure="Wang", combine="BMA", verbose=FALSE)
gene_sims_cc=mgeneSim(rownames(interesting_genes), semData=hsGO_cc, measure="Wang", combine="BMA", verbose=FALSE)

gene_sims=mgeneSim(rownames(interesting_genes), semData=hsGO2, measure="Wang", combine="BMA", verbose=FALSE)
gene_matrix_old=gene_matrix
gene_matrix=gene_sims

gene_matrix[is.na(gene_matrix)] <- 0

 # mean(gene_matrix, na.rm=T)
 # 
 # mean(gene_matrix[lista1_bm$ensembl_gene_id,lista1_bm$ensembl_gene_id], na.rm=T)
 # 
 # 
for(i in 1:dim(gene_matrix)[1]){
  for(j in 1:dim(gene_matrix)[2]){
    if(i==j){
      gene_matrix[i,j]=NA
    }
    if(i<j){
      gene_matrix[i,j]=NA
    }
  }
}

 
 

gene_matrix_mf=gene_sims_mf
gene_matrix_cc=gene_sims_cc



 gene_sims_mf[is.na(gene_sims_mf)] <- 0


for(i in 1:dim(gene_sims_mf)[1]){
  for(j in 1:dim(gene_sims_mf)[2]){
    if(i==j){
      gene_sims_mf[i,j]=NA
    }
    if(i<j){
      gene_sims_mf[i,j]=NA
    }
  }
}
 
# mean(gene_sims_mf, na.rm=T)
 
 
# mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id)], na.rm=T)
# mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id)], na.rm=T)
# mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id)], na.rm=T)
# 

 
 
 
  gene_sims_cc[is.na(gene_sims_cc)] <- 0


for(i in 1:dim(gene_sims_cc)[1]){
  for(j in 1:dim(gene_sims_cc)[2]){
    if(i==j){
      gene_sims_cc[i,j]=NA
    }
    if(i<j){
      gene_sims_cc[i,j]=NA
    }
  }
}

   # mean(gene_sims_cc, na.rm=T)
 
 
 # mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id)], na.rm=T)
 #  mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id)], na.rm=T)
 #    mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id)], na.rm=T)
 #    
  
lista.mod.copy2_id=lista.mod.copy

for(i in 1:length(lista.mod.copy2_id)){
lista.mod.copy2_id[[i]] <- bitr(lista.mod.copy[[i]], fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)$ENTREZID
}
 #     
 mat_emtrez=DOSE::geneSim(lista.mod.copy2_id$all,  measure="Wang", combine="BMA")
 mat_emtrez[is.na(mat_emtrez)] <- 0
 
 
 for(i in 1:dim(mat_emtrez)[1]){
  for(j in 1:dim(mat_emtrez)[2]){
    if(i==j){
      mat_emtrez[i,j]=NA
    }
    if(i<j){
      mat_emtrez[i,j]=NA
    }
  }
}

# mean(mat_emtrez, na.rm=T)
# mean(mat_emtrez[lista_cluster_ENTREZ$one,lista_cluster_ENTREZ$one], na.rm=T)
# mean(mat_emtrez[lista_cluster_ENTREZ$two,lista_cluster_ENTREZ$two], na.rm=T)
# mean(mat_emtrez[lista_cluster_ENTREZ$three,lista_cluster_ENTREZ$three], na.rm=T)
# 

 

 

# 
# gs1 <- c("835", "5261","241", "994", "514", "533")
# gs2 <- c("578","582", "400", "409", "411")
# clusterSim(gs1, gs2, semData=hsGO, measure="Wang", combine="BMA")
# 
# library(org.Hs.eg.db)
# x <- org.Hs.egGO
# hsEG <- mappedkeys(x)
# set.seed <- 123
# clusters <- list(a=sample(hsEG, 20), b=sample(hsEG, 20), c=sample(hsEG, 20))
# mclusterSim(clusters, semData=hsGO, measure="Wang", combine="BMA")
library(clusterProfiler)
lista_cluster=list(lista1_bm$ensembl_gene_id, lista2_bm$ensembl_gene_id,lista3_bm$ensembl_gene_id)
lista_cluster_ENTREZ=lista_cluster
for(i in 1:length(lista_cluster)){
lista_cluster_ENTREZ[[i]] <- bitr(lista_cluster[[i]], fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)$ENTREZID
}

names(lista_cluster_ENTREZ)=c("one", "two", "three")


 nperm=1000
list_perm_bp=rep(NA, nperm)
list_perm_mf=rep(NA, nperm)
list_perm_cc=rep(NA, nperm)

list_perm_do=rep(NA, nperm)

list_perm_bp2=rep(NA, nperm)
list_perm_mf2=rep(NA, nperm)
list_perm_cc2=rep(NA, nperm)

list_perm_do2=rep(NA, nperm)

 set.seed(1)
 for(i in 1:nperm){
   permuted=sample(lista.mod.copy2_id$all, length(lista_cluster_ENTREZ$one))
   permuted2=sample(lista.mod.copy2_id$all, length(lista_cluster_ENTREZ$three))

  
  permuted_ens=sample(rownames(interesting_genes), length(lista_cluster_ENTREZ$one))
  permuted_ens2=sample(rownames(interesting_genes), length(lista_cluster_ENTREZ$three))

     list_perm_bp[i]=mean(gene_matrix[intersect(rownames(gene_matrix),permuted_ens),intersect(rownames(gene_matrix),permuted_ens)], na.rm=T)
    list_perm_bp2[i]=mean(gene_matrix[intersect(rownames(gene_matrix),permuted_ens2),intersect(rownames(gene_matrix),permuted_ens2)], na.rm=T)
    
    list_perm_mf[i]=mean(gene_sims_mf[intersect(rownames(gene_sims_mf),permuted_ens),intersect(rownames(gene_sims_mf),permuted_ens)], na.rm=T)
    list_perm_mf2[i]=mean(gene_sims_mf[intersect(rownames(gene_sims_mf),permuted_ens2),intersect(rownames(gene_sims_mf),permuted_ens2)], na.rm=T)
    
    list_perm_cc[i]=mean(gene_sims_cc[intersect(rownames(gene_sims_cc),permuted_ens),intersect(rownames(gene_sims_cc),permuted_ens)], na.rm=T)
    list_perm_cc2[i]=mean(gene_sims_cc[intersect(rownames(gene_sims_cc),permuted_ens2),intersect(rownames(gene_sims_cc),permuted_ens2)], na.rm=T)
    
    list_perm_do[i]=mean(mat_emtrez[intersect(rownames(mat_emtrez),permuted),intersect(rownames(mat_emtrez),permuted)], na.rm=T)
    list_perm_do2[i]=mean(mat_emtrez[intersect(rownames(mat_emtrez),permuted2),intersect(rownames(mat_emtrez),permuted2)], na.rm=T)
 }
 
 
  
#  
#  > (sum(list_perm>mean(mat_emtrez[lista_cluster_ENTREZ$one,lista_cluster_ENTREZ$one], na.rm=T))+1)/(nperm+1)
# [1] 0.006993007
# > (sum(list_perm>mean(mat_emtrez[lista_cluster_ENTREZ$two,lista_cluster_ENTREZ$two], na.rm=T))+1)/(nperm+1)
# [1] 0.8841159
# > (sum(list_perm>mean(mat_emtrez[lista_cluster_ENTREZ$three,lista_cluster_ENTREZ$three], na.rm=T))+1)/(nperm+1)
# [1] 0.7762238
#  
#  find_permutation=function(list_perm, mean_value){
#    (sum(list_perm>mean_value)+1)/(nperm+1)
#  }
# 
# 
# find_permutation(list_perm_bp, mean(gene_matrix[intersect(rownames(gene_matrix),lista1_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista1_bm$ensembl_gene_id)], na.rm=T))
# 
# find_permutation(list_perm_bp, mean(gene_matrix[intersect(rownames(gene_matrix),lista2_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista2_bm$ensembl_gene_id)], na.rm=T))
# find_permutation(list_perm_bp2, mean(gene_matrix[intersect(rownames(gene_matrix),lista3_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista3_bm$ensembl_gene_id)], na.rm=T))
# # 0.002997003
# 
# find_permutation(list_perm_mf, mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id)], na.rm=T))
# 
# find_permutation(list_perm_mf, mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id)], na.rm=T))
# find_permutation(list_perm_mf2, mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id)], na.rm=T))
# 
# find_permutation(list_perm_cc, mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id)], na.rm=T))
# 
# find_permutation(list_perm_cc, mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id)], na.rm=T))
# # 0.003996004
# find_permutation(list_perm_cc2, mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id)], na.rm=T))
# # 0.04395604
# 
#  
# 
# find_permutation(list_perm_do,mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$one),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$one)], na.rm=T))
# # 0.001998002
# find_permutation(list_perm_do,mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$two),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$two)], na.rm=T))
# find_permutation(list_perm_do2,mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$three),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$three)], na.rm=T))
# 


```

```{r, fig.height=20}
par(mfcol=c(4,1))
boxplot(list("permutations"=list_perm_bp, "cluster 1"=mean(gene_matrix[intersect(rownames(gene_matrix),lista1_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista1_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 2"=mean(gene_matrix[intersect(rownames(gene_matrix),lista2_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista2_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 3"=mean(gene_matrix[intersect(rownames(gene_matrix),lista3_bm$ensembl_gene_id),intersect(rownames(gene_matrix),lista3_bm$ensembl_gene_id)], na.rm=T)), main="mean similarity (BP)")

boxplot(list("permutations"=list_perm_mf, "cluster 1"=mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista1_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 2"=mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista2_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 3"=mean(gene_sims_mf[intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_mf),lista3_bm$ensembl_gene_id)], na.rm=T)), main="mean similarity (MF)")

boxplot(list("permutations"=list_perm_cc, "cluster 1"=mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista1_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 2"=mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista2_bm$ensembl_gene_id)], na.rm=T),
                           "cluster 3"=mean(gene_sims_cc[intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id),intersect(rownames(gene_sims_cc),lista3_bm$ensembl_gene_id)], na.rm=T)), main="mean similarity (CC)")

boxplot(list("permutations"=list_perm_do, "cluster 1"=mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$one),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$one)], na.rm=T),
                          "cluster 2"= mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$two),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$two)], na.rm=T),
                           "cluster 3"=mean(mat_emtrez[intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$three),intersect(rownames(mat_emtrez),lista_cluster_ENTREZ$three)], na.rm=T)), main="mean similarity (DO)")

```


## Checking with human T-cell dataset

SRP078912

```{r}
pthreshold=.05
load("~/OneDrive - CRG - Centre de Regulacio Genomica/Dropbox (CRG)/metanalysis/SRP078912_tcells.RData")
summary(res)
resSRP078912_l=res
universe_human=rownames(resSRP078912_l)
resSRP078912_l=as.data.frame(resSRP078912_l)
resSRP078912_l=resSRP078912_l[order(resSRP078912_l$padj),]
resSRP078912_l=resSRP078912_l[complete.cases(resSRP078912_l),]
resSRP078912_l_sig=resSRP078912_l[which(resSRP078912_l$padj<pthreshold),]

resSRP078912_l_sigs=rownames(resSRP078912_l_sig)[1:500]
```

```{r}
count_table<-function(category, universe, de_list_inuniverse, de_list_incat){#given number of element in each list create a contigency table for performing fisher test
  mat=matrix(c(de_list_incat, 
               de_list_inuniverse-de_list_incat,
               category-de_list_incat,
               universe-(category-de_list_incat+de_list_inuniverse)),2)
}

category=length(intersect(universe_human, rownames(interesting_genes)))
universe=length(universe_human)
de_list_inuniverse=length( interesting_genes$ensembl_gene_id)
de_list_incat=length(intersect(rownames(interesting_genes), resSRP078912_l_sigs))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")

#14%overlap

phyper(de_list_incat-1, de_list_inuniverse, universe-de_list_inuniverse, category,lower.tail= FALSE)
# phyper(Overlap-1, group2, Total-group2, group1,lower.tail= FALSE)
#9.288694e-63

library(eulerr)
plot(euler(list("consistently DE genes"=rownames(interesting_genes),"T-cells (GSE84531)"=resSRP078912_l_sigs), shape = "ellipse"), quantities = TRUE)



par(mar=c(4,8,2,2))
barplot(c(length(intersect(resSRP078912_l_sigs, DS_brain_studies)),
length(intersect(resSRP078912_l_sigs, DS_blood_studies)),
length(intersect(resSRP078912_l_sigs, DS_placenta_amnios_studies)),
length(intersect(resSRP078912_l_sigs, DS_undifferentiated_studies)),
length(intersect(resSRP078912_l_sigs, DS_fibroblasts_studies)),
length(intersect(resSRP078912_l_sigs, entropicgenes))),
names.arg=c("brain", "blood", "placenta/amnios", "iPSCs/ESCs", "fibroblasts", "entropic genes"), xlab="number of genes",
col=c(igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="brain")],
              igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="blood")],
              igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="placenta/amnios")],
              igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="iPSCs/ESCs")],
              igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="fibroblasts")],"black"
              ), horiz=T, las=2)


# 
# par(mar=c(4,8,2,2))
# barplot(c(length(intersect(resSRP078912_l_sigs, DS_brain_studies))/length(DS_brain_studies)*100,
# length(intersect(resSRP078912_l_sigs, DS_blood_studies))/length(DS_blood_studies)*100,
# length(intersect(resSRP078912_l_sigs, DS_placenta_amnios_studies))/length(DS_placenta_amnios_studies)*100,
# length(intersect(resSRP078912_l_sigs, DS_undifferentiated_studies))/length(DS_undifferentiated_studies)*100,
# length(intersect(resSRP078912_l_sigs, DS_fibroblasts_studies))/length(DS_fibroblasts_studies)*100,
# length(intersect(resSRP078912_l_sigs, entropicgenes))/length(entropicgenes)*100),
# names.arg=c("brain", "blood", "placenta/amnios", "iPSCs/ESCs", "fibroblasts", "entropic genes"), xlab="% of genes",
# col=c(igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="brain")],
#               igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="blood")],
#               igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="placenta/amnios")],
#               igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="iPSCs/ESCs")],
#               igraph::V(graph_coex2)$color[which(igraph::V(graph_coex2)$name =="fibroblasts")],"black"
#               ), horiz=T, las=2)


```

### Summary interesting genes

```{r}
library(eulerr)
set.seed(1)
plot(euler(list("consistently DE"=interesting_genes$ensembl_gene_id,"consistently DE only in one tissue macro-category"=setdiff(unique(unlist(DS_tissue_genes)), entropicgenes), "consistently DE in more tissue macro-categories"=entropicgenes, "consistently upregulated/ downregulated"=highest_changing_genes),shape = "ellipse"), quantities = TRUE)

```
```{r}
summary_forapp=summary[,c("GEO.ArrayExpress", "SRA_study", "Species" ,"Model","Platform","Cell_type","Tissue","PMID", "Design", "Description"      ,"Notes" )]
whiteblue=colorRampPalette(c("white", "blue"))
cutcolors=whiteblue(13)[gene_matrix_old[V(graph_coex)$name,3]]

```


```{r}
save.image("meta_analysis.rda")
# save(summary_forapp, allgenes, list_array, interesting_genes,cutcolors, edge_color, graph_coex, mod.copy_coex, file="forapp.rda")

write.csv(interesting_genes, "coDEgenes_microcategories.csv")
write.csv(interesting_genes_macro, "coDEgenes_macrocategories.csv")

```


```{r}
# library(easyPubMed)
# summary=read.csv("summary_metanalysis_DS.csv",  header=T)
# pmid=unique(summary$PMID)
# pmid=pmid[!is.na(pmid)]
# 
# data_papers=matrix(length(pmid)*4, ncol=4, nrow=length(pmid))
# colnames(data_papers)=c("PMID", "YEAR", "MONTH", "DAY")
# for(i in 1:length(pmid)){
# my_entrez_id <- get_pubmed_ids(pubmed_query_string=as.character(pmid[i]))
# fetched <- fetch_pubmed_data(my_entrez_id)
# data_papers[i, 1]=unique(article_to_df(fetched)$pmid)
# data_papers[i, 2]=unique(article_to_df(fetched)$year)
# data_papers[i, 3]=unique(article_to_df(fetched)$month)
# data_papers[i, 4]=unique(article_to_df(fetched)$day)
#   }
# 
# data_papers=as.data.frame(data_papers)
# data_papers$YEAR=as.numeric(data_papers$YEAR)
# data_papers$MONTH=as.numeric(data_papers$MONTH)
# data_papers$DAY=as.numeric(data_papers$DAY)
# 
# data_papers=data_papers[order(data_papers$YEAR, data_papers$MONTH,data_papers$DAY),]
# Last paper used in the metanalysis the `r data_papers[length(pmid),2]`/`r data_papers[length(pmid),3]`/`r data_papers[length(pmid),4]`
```

#### Venn diagram with Vilardell data

```{r venn}
library(eulerr)

plot(euler(list("Vilardell"=previous$Ensembl, "Consistently DE"=interesting_genes$ensembl_gene_id), shape = "ellipse"), quantities = TRUE)


```

```{r}
count_table<-function(category, universe, de_list_inuniverse, de_list_incat){#given number of element in each list create a contigency table for performing fisher test
  mat=matrix(c(de_list_incat, 
               de_list_inuniverse-de_list_incat,
               category-de_list_incat,
               universe-(category-de_list_incat+de_list_inuniverse)),2)
}

category=length(previous$Ensembl)
universe=length(unique(unlist(list_ensembl)))
de_list_inuniverse=length( interesting_genes$ensembl_gene_id)
de_list_incat=length(intersect(previous$Ensembl, interesting_genes$ensembl_gene_id))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")

```

```{r}
mirna_hsa21=c("hsa-mir-8069-1","hsa-mir-6274-1","hsa-mir-10401","hsa-mir-3648-1","hsa-mir-6274-2","hsa-mir-6274-3","hsa-mir-6274-4","hsa-mir-10396b","hsa-mir-3648-2","hsa-mir-10396a","hsa-mir-3156-3","hsa-mir-3118-1","hsa-mir-8069-2","hsa-mir-99a","hsa-mir-let-7c","hsa-mir-125b-2","hsa-mir-548x","hsa-mir-6130","hsa-mir-155","hsa-mir-4759","hsa-mir-4327","hsa-mir-6501","hsa-mir-802","hsa-mir-6508","hsa-mir-4760","hsa-mir-3197","hsa-mir-6814","hsa-mir-5692b","hsa-mir-6070","hsa-mir-6815")

library(multiMiR)


multimir_results <- get_multimir(org     = 'hsa',
                                 mirna   = mirna_hsa21,
                                 table   = 'validated',
                                 summary = TRUE)

targets=unique(multimir_results@data$target_ensembl)

intersect(targets,interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC>0)])

intersect(targets,interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC<0)])

length(intersect(targets,interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC>0)]))/length(interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC>0)])

length(intersect(targets,interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC<0)]))/length(interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC<0)])

upgenes=interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC>0)]

category=length(intersect(universe_human, targets))
universe=length(universe_human)
de_list_inuniverse=length( intersect(upgenes, universe_human))
de_list_incat=length(intersect(intersect(upgenes, universe_human), intersect(universe_human, targets)))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")
#p-value = 4.734e-07

downgenes=interesting_genes$ensembl_gene_id[which(interesting_genes$mean_log2FC<0)]

category=length(intersect(universe_human, targets))
universe=length(universe_human)
de_list_inuniverse=length( intersect(downgenes, universe_human))
de_list_incat=length(intersect(intersect(downgenes, universe_human), intersect(universe_human, targets)))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")

#p-value = 7.822e-06
```

```{r}
upgenescons=highest_changing_genes_table$ensembl_gene_id[which(highest_changing_genes_table$mean_log2FC>0)]

category=length(intersect(universe_human, targets))
universe=length(universe_human)
de_list_inuniverse=length( intersect(upgenescons, universe_human))
de_list_incat=length(intersect(intersect(upgenescons, universe_human), intersect(universe_human, targets)))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")
#p-value = 0.0374

downgenescons=highest_changing_genes_table$ensembl_gene_id[which(highest_changing_genes_table$mean_log2FC<0)]

category=length(intersect(universe_human, targets))
universe=length(universe_human)
de_list_inuniverse=length( intersect(downgenescons, universe_human))
de_list_incat=length(intersect(intersect(downgenescons, universe_human), intersect(universe_human, targets)))
mat=count_table(category, universe, de_list_inuniverse, de_list_incat)
fisher.test((mat), alternative="greater")

#p-value = 0.001044
```
consistently downregulated genes that are target of mi-R mapping on HSA21

```{r}
Table=multimir_results@data[which( multimir_results@data$target_ensembl %in% intersect(intersect(downgenescons, universe_human), intersect(universe_human, targets))),]


kable(Table, "html") %>%
    kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = T) 

```


